---
output:
  word_document: default
  html_document: default
---
# Map Unit GIS Summary Report

This report summarizes the geographic setting of a list of musym within a file geodatabase. It is intented to be used to compare and contrast map units, and suggest possible Low, RV, and High values for components. 

### Units
Elevation = meters,
Slope = percent,
Aspect = degrees,
Valley = unitless,
Wetness = unitless,
Relief = meters,
Ppt = millimeters,
Temp = Celsius,
Ffp = days,
Circularity is a ratio of polygon length / circumference of the polygon were it a perfect circle. So the circularity of a perfect circle = 1. This is an estimate of shape complexity.


```{r LoadPackages, include=F}
library(circular)
library(gdalUtils)
library(sp)
library(raster)
library(rgdal)
library(reshape)
library(plyr)
library(lattice)
library(xtable)
```

```{r load-data, echo=FALSE, results='hide', message=FALSE}
mukey <- c('159669','159758','160795','160824','162087','163213','164424','1612865','1613014') # List of mukey
dir <- "M:/geodata/project_data/"
geodatabase <- "RTSD_Region_11_IND_FY15.gdb"
office <- "11IND"

# Defaults
setwd("C:/R/stools/mapunitSummary")
gdal_setInstallation(search_path="C:/Program Files/QGIS Dufour/bin", rescan=T)
samplePercent <- 10
p <- c(0,0.25,0.5,0.75,1)

pd <- paste0(dir, office, "/")
rd <- paste0(dir, "11REGION/")
shp <- paste("mu", noquote(paste(mukey, collapse=".", sep="")), sep="")
cache <- paste0(pd, "cache/", shp, ".shp")

# Load packages

## local functions
mapunit.summary<-function(mukey, samplePercent){
  # Import geodatabase
  if(ifelse(file.exists(cache), "NA", 1) == 1) {
   ogr2ogr(
      src_datasource_name=paste0(pd, geodatabase),
      dst_datasource_name=cache,
      layer="MUPOLYGON",
      where=paste("MUKEY IN (", noquote(paste("'", mukey, "'", collapse=",", sep="")),")", sep=""),
      s_srs=CRS("+init=epsg:5070"),
      t_srs=CRS("+init=epsg:5070"),
      overwrite=T,
      verbose=TRUE
      )
    sapolygon <- readOGR(dsn=paste0(pd, geodatabase), layer="SAPOLYGON", encoding="OpenFileGDB")
    mupolygon <- readOGR(dsn=cache, layer=shp, encoding="ESRI Shapefile")
    mupolygon@data$Acres <- mupolygon@data$Shape_Area * 0.000247105
    mupolygon@data$Circularity <- mupolygon@data$Shape_Leng/(2*sqrt(mupolygon@data$Shape_Area/pi)*pi)
    n <- sum(data.frame(mupolygon)$Shape_Area)/900*samplePercent/100

    # Sample soil map units
    mupolygon.s<-spsample(mupolygon, n=n, "stratified")
    mapunit.df<-cbind(mupolygon.s,over(mupolygon.s,mupolygon))

    # Load grids
    g10.st <- stack(
      paste0(pd, "ned10m_", office, "_slope5.tif"), 
      paste0(pd, "ned10m_", office, "_aspect5.tif")
      )
    g30.st <- stack(
      paste0(pd, "ned30m_", office, ".tif"),
      paste0(pd, "ned30m_", office, "_wetness.tif"),
      paste0(pd, "ned30m_", office, "_mvalleys.tif"),
      paste0(pd, "ned30m_", office, "_z2stream.tif")
      )
    g800.st <- stack(
      paste0(rd, "prism800m_11R_ppt_1981_2010_annual_mm.tif"),
      paste0(rd, "prism800m_11R_tmean_1981_2010_annual_C.tif")
      )
    g1000.st <- stack(
      paste0(rd, "rmrs1000m_11R_ffp_1961_1990_annual_days.tif")
      )
  
    names(g10.st) <- c("slope", "aspect")
    names(g30.st) <- c("elev", "wetness", "valleys", "relief")
    names(g800.st) <- c("ppt", "temp")
    names(g1000.st) <- c("ffp")

  
    # Extract geodata
    g10.e <- extract(g10.st, mupolygon.s)
    g30.e <- extract(g30.st, mupolygon.s)
    g800.e <- extract(g800.st, mupolygon.s)
    g1000.e <- extract(g1000.st, mupolygon.s)
    geodata.df <- data.frame(g10.e, g30.e, g800.e, g1000.e)
  
    # Prep data
    data <- cbind(mapunit.df$MUKEY, geodata.df)
    names(data)[1] <- "MUKEY"
    data$aspect <- circular(data$aspect, template="geographic", units="degrees", modulo="2pi")
  
    slope.list<-c(0,2,4,8,15,30,50,75,350)
    aspect.list<-c(0,23,68,113,158,203,248,293,338,360) 
    valleys.list<-c(0,0.5,30)
  
    data$slopeBreaks <- cut(data$slope,breaks=slope.list, right=FALSE)
    data$aspectBreaks <- cut(data$aspect,breaks=aspect.list, right=FALSE)
    data$valleysBreaks <- cut(data$valleys,breaks=valleys.list, right=FALSE)

    levels(data$slopeBreaks) <- c("0-2","2-4","4-8","8-15","15-30","30-50","50-75","75-350")
    levels(data$aspectBreaks) <- c("N","NE","E","SE","S","SW","W","NW","N")
    levels(data$valleysBreaks) <- c("upland","lowland")
  
  data.l <- list(data=data, mupolygon=mupolygon, sapolygon=sapolygon)
  save(data.l, file=paste0(cache, ".Rdata"))
  return(data.l)
  } else load(file=paste0(cache, ".Rdata")) #loading
  return(data.l)
}

conditional.l.rv.h.summary <- function(x) {
  variable <- unique(x$variable)
  v <- na.omit(x$value) # extract column, from long-formatted input data
  precision <- if(variable == 'Circularity') 1 else 0
  ci <- quantile(v, na.rm=TRUE, probs=p) 
  d <- data.frame(min=ci[1], low=ci[2], rv=ci[3], high=ci[4], max=ci[5], stringsAsFactors=FALSE) # combine into DF
  d$range <- with(d, paste("(", paste(round(c(min, low, rv, high, max), precision),collapse=', '), ")", sep="")) # add 'range' column for pretty-printing
  return(d[6])
}

data.l <- mapunit.summary(mukey, samplePercent)
data <- data.l$data
mupolygon <- data.l$mupolygon
sapolygon <- data.l$sapolygon
h <- length(levels(data.l$MUKEY))*3

data2 <- data
data2$MUKEY <- "representative"
data <- rbind(data, data2)
```

## Polygon metrics
```{r echo=FALSE, results='asis'}
pol <- data.frame(mupolygon)
pol.m <- melt(pol, id.vars="MUKEY", measure.vars=c("Acres", "Circularity"))
pol.cs <- ddply(pol.m, .(variable, MUKEY), .fun=conditional.l.rv.h.summary)
pol.n <- ddply(pol, .(MUKEY), .drop=T, summarize, nPolygons=length(MUKEY), nAcres=sum(Acres))
pol.cs.c <-cast(pol.cs, MUKEY ~ variable, value='range')
pol.me <- merge(pol.cs.c, pol.n, by="MUKEY")
print(xtable(pol.me, caption="Five number summary (e.g. min, 25th, median, 75th, max)", digits=0), type="html", caption.placement="top")

pol.as.m <- melt(pol, id.vars=c("MUKEY", "AREASYMBOL"), measure.vars=c("Acres"))
pol.as.cs1 <- ddply(pol.as.m, .(MUKEY, AREASYMBOL, variable), .fun=conditional.l.rv.h.summary)
pol.as.cs2 <- ddply(pol.as.m, .(MUKEY, AREASYMBOL, variable), summarize, nPolygons=length(value), nArces=sum(value))[, c(1,2,4,5)]
pol.as.me <- merge(pol.as.cs1, pol.as.cs2, by=c("MUKEY", "AREASYMBOL"))
print(xtable(pol.as.me, digits=0, caption="Summary of MUKEY per AREASYMBOL"), type="html", caption.placement="top")
```

## Polygon Boxplots
### *(min, 5th, 25th, median, 75th, 95th, max)*
```{r echo=FALSE}
bwplot(MUKEY ~ value|variable, data=pol.m, scales=list(x="free"))
```

## Percentages
```{r echo=FALSE, results='asis'}
## Create descriptive and graphical summary of map unit
print(xtable(prop.table(xtabs(~MUKEY+slopeBreaks,data=data), margin=1)*100, digits=0, caption="Slope breaks"), type="html", caption.placement="top")
print(xtable(prop.table(xtabs(~MUKEY+aspectBreaks,data=data), margin=1)*100, digits=0, caption="Aspect breaks"), type="html", caption.placement="top")

print(xtable(prop.table(xtabs(~MUKEY+valleysBreaks,data=data), margin=1)*100, digits=0, caption="Upland vs. lowland"), type="html", caption.placement="top")
```

## Percentiles
### *(e.g. min, 25th, median, 75th, max)*
```{r, echo=FALSE, results="asis"}
data.m <- melt(data, id.vars="MUKEY", measure.vars=c("elev", "slope", "valleys", "wetness", "relief", "ppt", "temp", "ffp"))
data.cs <- ddply(data.m, .(variable,MUKEY), .fun=conditional.l.rv.h.summary)
data.cs.c <- cast(data.cs, MUKEY ~ variable, value='range')
data.n <- ddply(data, .(MUKEY), .drop=T, summarize, n=length(MUKEY))
print(xtable(merge(data.cs.c, data.n, by="MUKEY")), type="html")
aspect.m <- melt(data, id.vars="MUKEY", measure.vars=c("aspect"))
aspect.m$value <- circular(aspect.m$value, template="geographic", units="degrees", modulo="2pi")
aspect.cs <- ddply(aspect.m, .(variable,MUKEY), .fun=conditional.l.rv.h.summary)
aspect.cs.c <- cast(aspect.cs, MUKEY ~ variable, value='range')
print(xtable(merge(aspect.cs.c, data.n, by="MUKEY")), type="html")
```

## Plot of soil map units
```{r echo=FALSE}
spplot(mupolygon, zcol="MUKEY", scales=list(draw=T), col.regions=bpy.colors(length(mukey)), lty=0, sp.layout=list("sp.polygons", sapolygon))
```

## Boxplots
### *(min, 5th, 25th, median, 75th, 95th, max)*
```{r echo=FALSE}
bwplot(MUKEY ~ value|variable, data=data.m, scales=list(x="free"))          
```

