---
output:
  html_document: default
---
# Pedon report

```{r}
# Set Global Parameters
series <- "Genesee"
p <- c(0, 0.10, 0.5, 0.90, 1)
```

```{r LoadPackages, include=FALSE}
# load libraries
library(aqp)
library(soilDB)
library(reshape)
library(plyr)
library(lattice)
library(xtable)
```

```{r, load-data, echo=FALSE, warning=FALSE}
# source custom R functions
source("C:/Users/Stephen/Documents/soilRwb/pedonReport/genhz_rules.R")

# load NASIS data
f <- fetchNASIS()
f <- subsetProfiles(f, s="taxonname == series")
f@horizons$hzname[is.na(f@horizons$hzname)] <- "NA"
f@horizons$genhz <- generalize.hz(f@horizons$hzname, ghr$genesee$n, ghr$genesee$p)
names(f@horizons)[15] <- "fragvoltotc"
names(f@horizons)[16] <- "fragvoltot"

d_rgb <- f@horizons[, c("d_r", "d_g", "d_b")]
d_munsell <- rgb2munsell(d_rgb)
m_rgb <- f@horizons[, c("m_r", "m_g", "m_b")]
m_munsell <- rgb2munsell(m_rgb)

f@horizons$d_hue <- as.character(d_munsell$hue)
f@horizons$d_value <- as.integer(d_munsell$value)
f@horizons$d_chroma <- as.integer(d_munsell$chroma)
f@horizons$m_hue <- as.character(m_munsell$hue)
f@horizons$m_value <- as.integer(m_munsell$value)
f@horizons$m_chroma <- as.integer(d_munsell$chroma)

naReplace <- function(h){
  l <- list()
  for(i in seq(names(h))){
    if(class(h[,i])=="character") {l[[i]] <- replace(h[,i], is.na(h[,i]), "NA")} else(l[[i]] <-  h[,i])
  }
  l <- data.frame(l, stringsAsFactors=FALSE)
  names(l) <- names(h)
  return(l)
}

h2 <- naReplace(f@horizons)
f@horizons <- h2

h <- horizons(f)
s <- site(f)

# Function
conditional.l.rv.h.summary <- function(x) {
  variable <- unique(x$variable)
  v <- na.omit(x$value) # extract column, from long-formatted input data
  precision <- if(variable == 'phfield') 1 else 0
  ci <- quantile(v, na.rm=TRUE, probs=p) 
  d <- data.frame(min=ci[1], low=ci[2], rv=ci[3], high=ci[4], max=ci[5], stringsAsFactors=FALSE) # combine into DF
  d$range <- with(d, paste("(", paste(round(c(min, low, rv, high, max), precision), collapse=', '), ")", sep="")) # add 'range' column for pretty-printing
  return(d[6])
}
```
## Plot of all pedons in selected set
```{r}
plot(f, label="pedon_id")
```


# Summary of Sites
```{r, echo=FALSE}
# Site information
subset(s, select=c("pedon_id", "taxonname", "tax_subgroup", "part_size_class", "pedon_type", "describer"))

# Summarize surface rock fragments
s$total_srf <- s$surface_gravel+s$surface_cobbles+s$surface_stones+s$surface_boulders+s$surface_flagstones+s$surface_channers
s$gravel <- s$surface_fgravel-s$surface_gravel
s.m <- melt(s, measure.vars=c("total_srf", "surface_fgravel", "surface_gravel", "surface_cobbles", "surface_stones", "surface_boulders", "surface_channers", "surface_flagstones"))
s.summary <- ddply(s.m, .(variable), .fun=conditional.l.rv.h.summary)
```


## Box plots of surface rock fragments
```{r}
bwplot(variable~value, data=s.m, scales=list(x="free"), xlab="percent")          
```


## Range in characteristics of surface rock fragments (min, 10th, median, 90th, max)
```{r, results='asis'}
print(xtable(cast(s.summary, ~variable, value='range')), type="html")
```

# Summary of Pedons
```{r, echo=FALSE, results='asis'}
# Summarize numeric variables by generic horizon
h$gravel <- h$gravel-h$fine_gravel

hp.m <- melt(h, id.vars="genhz", measure.vars=c('clay', 'sand', 'fine_gravel', 'gravel', 'cobbles', 'stones', 'fragvoltot'))
hp.cs <- ddply(hp.m, .(variable, genhz), .fun=conditional.l.rv.h.summary)

hc.m <- melt(h, id.vars="genhz", measure.vars=c('phfield', 'd_value', 'd_chroma', 'm_value', 'm_chroma'))
hc.cs <- ddply(hc.m, .(variable, genhz), .fun=conditional.l.rv.h.summary)

thk <- ddply(h, .(peiid, genhz), summarize, thickness=sum(hzdepb-hzdept))
thk.m <- melt(thk, id.vars="genhz", measure.vars="thickness")
thk.cs <- ddply(thk.m, .(variable, genhz), .fun=conditional.l.rv.h.summary)

hp.c <- cast(hp.cs, genhz ~ variable, value='range')
hc.c <- cast(hc.cs, genhz ~ variable, value='range')
```


## Horizon designations by generic horizon
```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(f@horizons$genhz, f@horizons$hzname)), digits=0), type="html") 
```


## Range in characteristics for generic horizons (min, 10th, median, 90th, max)
```{r, results='asis'}
print(xtable(hp.c), type="html")
print(xtable(hc.c), type="html")
```

## Box plots of numeric variables by generic horizon
```{r}
bwplot(genhz ~ value|variable, data=rbind(hp.m, hc.m, thk.m), scales=list(x="free"))          
```

## Texture by generic horizon
```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$texture_class)), digits=0), type="html")
```

## Tex Mod & Class by generic horizon
```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$texture)), digits=0), type="html")
```

## Dry hue by generic horizon
```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$d_hue)), digits=0), type="html")
```

## Moist hue by generic horizon
```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$m_hue)), digits=0), type="html")
```

## Effervescence by generic horizon
```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$effervescence)), digits=0), type="html")
```


## Thickness of generic horizons
```{r, echo=FALSE, results='asis'}
genhz.thk <- ddply(h, .(phiid, genhz), summarize, thickness=sum(hzdepb-hzdept))
genhz.m <- melt(h, id.vars="genhz", measure.vars=c('hzdept', 'hzdepb'))
genhz.thk.m <- melt(genhz.thk, id.vars="genhz", measure.vars='thickness')
genhz.m <- rbind(genhz.m, genhz.thk.m)
d.cs <- ddply(genhz.m, .(variable, genhz), .fun=conditional.l.rv.h.summary)
print(xtable(cast(d.cs, genhz ~ variable, value='range'), digits=0), type="html")
```


## Boxplot of generic horizon thicknesses
```{r}
bwplot(genhz ~ value|variable, data=genhz.m, scales=list(x="free"), xlab="cm", horizontal=T)          
```

## Thickness of diagnostic horizons
```{r, echo=FALSE, results='asis'}
diag.thk <- ddply(f@diagnostic, .(peiid, diag_kind), summarize, thickness=sum(featdepb-featdept))
diag.m <- melt(f@diagnostic, id.vars="diag_kind", measure.vars=c('featdept', 'featdepb'))
diag.thk.m <- melt(diag.thk, id.vars="diag_kind", measure.vars='thickness')
diag.m <- rbind(diag.m, diag.thk.m)
d.cs <- ddply(diag.m, .(variable, diag_kind), .fun=conditional.l.rv.h.summary)
print(xtable(cast(d.cs, diag_kind ~ variable, value='range'), digits=0), type="html")
```
## Boxplot of diagnostic horizon thicknesses
```{r}
bwplot(diag_kind ~ value|variable, data=diag.m, scales=list(x="free"), xlab="cm")          
```

