---
title: "POX-C Supplement"
author: "Wills et al"
date: "March 12, 2018"
output: html_document
---

```{r setup, include=FALSE}

#rm(list=ls()) #clear previous data

require("knitr")
knitr::opts_chunk$set(echo = TRUE, comment = "#", warning = FALSE, message = FALSE, error =FALSE, tidy.opts=list(width.cutoff=60),tidy=TRUE)

#checks for packages and downloads as necessary
list.of.packages <- c("tidyverse", "ggthemes", "RCurl", "gridExtra", "GGally", "clhs", "ggbeeswarm", "ggridges" ,"ggjoy", "viridis", "readxl")

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(tidyverse)
library(RCurl)
library(gridExtra)
library(GGally)
library(clhs)
library(ggjoy)
library(ggbeeswarm)
library(ggridges)
library(viridis)
library(readxl)

```

##POX-C Project Analysis

#### Load Data
This will download relevent data for this project from a NRCS file sharing site. 

```{r data_get}

#POX (AC) sample pool used to select samples for further analysis
AC_url <- 'https://nrcs.box.com/shared/static/bm9o3nj55rstbsvoloysid75pcxf3tlr.csv'
ifelse(!file.exists('ACsamplelist.csv'), download.file(AC_url, destfile = paste0('ACsamplelist.csv')), "File already downloaded")

#POX (AC) samples selected for further analysis
ACselect_url <- 'https://nrcs.box.com/shared/static/wll2xi6v73b1f9re6t7mlhpl8afha4oe.csv'
ifelse(!file.exists('ACsampleselect.csv'), download.file(ACselect_url, destfile = paste0('ACsampleselect.csv')), "File already downloaded")


#Newly analyzed data, for this project
POX_url <- 'https://nrcs.box.com/shared/static/t9antdnnt502sa1x8a7qu5r7n3z4yk7v.csv'
ifelse(!file.exists('pox_project_data.csv'), download.file(POX_url, destfile = paste0('pox_project_data.csv')), "File already downloaded")


#other information about project samples
POX_info_url <- 'https://nrcs.box.com/shared/static/7ke7l4ob2rmw3zoqf16pm6a5m9qpoo5y.csv'
ifelse(!file.exists('pox_project_info.csv'), download.file(POX_info_url, destfile = paste0('pox_project_info.csv')), "File already downloaded")



```

###Select samples
This portion of the anlysis cannot be replicated so it is not rerun each time.  It is included to show how the samples were selected from the available KSSL samples at the time of analysis.


```{r select}

#all KSSL samples with available POXC measurements, auxiliary data and available sample
AClist <- read.csv('ACsamplelist.csv')

AClist$taxonomic.order <- as.factor(toupper(AClist$taxonomic.order))

levels(AClist$taxonomic.order) #standardize categorical variable

#visualize duplicated samples - the values seem to change when they've been run more than once
D <- ggplot(data=AClist, aes(x=Measured.Active.Carbon, fill=Dupl)) + geom_density()


dd <- ggplot(data=AClist, aes(y=Measured.Active.Carbon, x=taxonomic.order, color=Dupl)) + geom_boxplot() + theme(axis.title.x = element_text(face="bold", colour="#990000", size=20),axis.text.x  = element_text(angle=90, vjust=0.5, size=16))

grid.arrange(D,dd)

tC <- ggplot(data=AClist, aes(x=Total.Carbon, y = Measured.Active.Carbon, color= taxonomic.order)) + geom_point()
tCt <- ggplot(data=AClist, aes(x=Total.Carbon, y = Measured.Active.Carbon, color= Dupl)) + geom_point()

grid.arrange(tC, tCt)

#include all variable including smp.id - which is assigned chrnologically as a proxy for time of analysis
res <- clhs(AClist, size = 30, iter = 1000, progress = FALSE, simple = FALSE)

plot(res, modes = "dens")

#output selected samples for futher analysis
samp <- res$sampled_data
#write.csv(samp, 'ACsamp_select.csv')

#use the actual selection list
samp <-read_csv('ACsampleselect.csv')

#evaluate selection more specifically

#Add a value to ID both sets
samp <- samp %>% mutate(SET = "Selected", smp.id = as.factor(smp.id)) 
AClist <- AClist %>% mutate(SET = "Sample Pool", smp.id = as.factor(smp.id))

sampAC <- bind_rows(samp,AClist) %>% rename(MGMT = MGMT_lable)

S <- ggplot(data=sampAC, aes(x=Measured.Active.Carbon)) + geom_density(aes(fill=SET), alpha = 0.5) + scale_x_continuous(limits = c(0,10000))
Scl <- ggplot(data=sampAC, aes(x=Clay)) + geom_density(aes(fill=SET), alpha = 0.5)


Sd <- ggplot(data=sampAC, aes(y=Measured.Active.Carbon, x=taxonomic.order)) + geom_boxplot(aes(fill = SET)) +  theme(axis.title.x = element_text(face="bold", colour="#990000", size=20),axis.text.x  = element_text(angle=90, vjust=0.5, size=12)) + scale_y_continuous(limits = c(0,10000))


Sc <- ggplot(data=sampAC, aes(y=Measured.Active.Carbon, x=horz.desgn.master)) + geom_boxplot(aes(fill = SET)) + theme(axis.title.x = element_text(face="bold", colour="#990000", size=20),axis.text.x  = element_text(angle=90, vjust=0.5, size=12)) + scale_y_continuous(limits = c(0,10000))

grid.arrange(S,Scl, Sd, Sc, ncol = 2)

rm(AClist, res, samp, sampAC, tC, tCt, Sd,Sc)

```

#### Prepare Meausred Data for analysis
For each of the selected samples, new analysis was done - 3 replicates for each of a facotrial combination:  4 weights (0.25, 0.5, 1, 2.5, 5 g) and 2 grind sizes ( <2mm and < 80 um). Data is loaded, cleaning steps for analysis are included. 

```{r data_prep}

POX_info <- read.csv('pox_project_info.csv')

POX <- read.csv("pox_project_data.csv")

#ID blank and control
cb <- c('1','168')
CB <- c('BLANK', 'CONTROL')

POX$smp.id <- as.factor(ifelse(POX$smp_id == 1, 'BLANK',
                     ifelse(POX$smp_id == 168, 'CONTROL', as.character(POX$smp_id))))

#evaluate blanks and checks
ggplot(POX[POX$smp.id == CB,], aes(y=Conc,x=as.factor(smp.id))) + geom_boxplot(alpha=0.5)

POX.cb.summ <- POX %>%
    filter(smp.id == CB)%>%
    group_by(smp.id)%>%
    summarize(  N.conc   = length(Conc),
                 mean.conc = mean(Conc, na.rm=TRUE),
                 sd.conc   = sd(Conc, na.rm=TRUE),
                 se.conc   = sd.conc / sqrt(N.conc),
                 CV.conc = sd.conc/mean.conc,
                 N.pox   = length(AC),
                 mean.pox = mean(AC, na.rm=TRUE),
                 sd.pox   = sd(AC, na.rm=TRUE),
                 se.pox   = sd.pox / sqrt(N.pox),
                 CV.pox= sd.pox/mean.pox
                 )
 
POX.cb.summ

#information about selected samples (remove blanks and controls)
POX <- POX[!(grepl("BLANK", POX$smp.id) | grepl("CONTROL", POX$smp.id)) | !(grepl("", POX$Validation)),]

POX <- droplevels.data.frame(POX)

#treat sample amount/weight as a class
POX$wt_class <- factor(POX$wt_class)

#treat preparation/grind as facors and relabel
POX$Grind <- factor(POX$Grind,
                     levels = c("lt2mm", "fine_lt80um"),
                     labels = c("< 2 mm Grind", "Fine Grind"))


```


####Eploratory Data Analysis
First, both the mean and standard deviation of each sample is calculated (from 3 replicates). Then a summary table is produced by Grind and Weight.

```{r summary, echo=T}
#summary by sample
POX.smp.summ <- POX %>%
    filter(smp.id != CB, !is.na(wt_class))%>%
    mutate (Weight = wt_class) %>%
    group_by(smp.id, Grind, Weight)%>%
    summarize(Clay = mean(Clay), Total.Carbon = mean(Total.Carbon), pH = mean(pH_H2O), 
              N.conc   = length(Conc),
              mean.conc = mean(Conc, na.rm=TRUE),
              sd.conc   = sd(Conc, na.rm=TRUE),
              se.conc   = sd.conc / sqrt(N.conc),
              CV.conc = sd.conc/mean.conc,
              N.pox   = length(AC),
              mean.pox = mean(AC, na.rm=TRUE),
              sd.pox   = sd(AC, na.rm=TRUE),
              se.pox   = sd.pox / sqrt(N.pox),
              CV.pox= sd.pox/mean.pox,
              Concentration = mean.conc,
              POX = mean.pox
                 )


#summary by amount and grind
POX.summ <- POX.smp.summ %>%
    group_by(Grind, Weight)%>%
    summarize(N.conc   = length(mean.conc),
              mean.conc = mean(mean.conc, na.rm=TRUE),
              avg.sd.conc =  mean(sd.conc, na.rm=TRUE),
              avg.se.conc   = mean(se.conc, na.rm=TRUE),
              avg.CV.conc = mean(CV.conc, na.rm=TRUE),
              N.pox   = length(mean.pox),
              mean.pox = mean(mean.pox, na.rm=TRUE),
              avg.sd.pox = mean(sd.pox, na.rm=TRUE),
              avg.se.pox = mean(se.pox, na.rm=TRUE),
              avg.CV.pox= mean(CV.pox, na.rm=TRUE),
              Concentration = mean.conc,
              POX = mean.pox
                 )

kable(POX.summ %>% select(Concentration, Weight, Grind) %>% spread(Weight, Concentration), digits = 5, align = 'c', caption = "Average Solution Concentration \n by Grind size and Analyzed Weight" )


kable(POX.summ %>% select(Grind, Weight, POX) %>% spread(Weight, POX), digits = 1, align = 'c', caption = "Average POX C \n by Grind size and Analyzed Weight" )

kable(POX.summ %>% select(Grind, Weight, avg.sd.pox) %>% spread(Weight, avg.sd.pox), digits = 2, align = 'c', caption = "SD of POX \n Average by Grind size and Analyzed Weight" )


#library(kableExtra)
# kable(POX.summ %>% select(Grind, Weight, Concentration, POX), "html") %>%
#   kable_styling(boostrap_options = c("striped", "hover"))

    
                
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r explore}

a <- POX %>% ggplot(aes(y = Conc, x = wt_class, color = Validation)) + geom_point(size = 6) + labs(x = "Sample Weight", y = 'Concetration', title = expression(paste('Pottasium Permanganate Analysis','KMnO'[4])))


b <- POX.smp.summ %>% ggplot(aes(y = POX, x = Weight, color = Grind)) + geom_boxplot(aes(fill = Grind)) +labs(x = 'Weight, g', y = expression(paste('POX C, ',' mg kg'^-1)), title = expression(paste('Pottasium Permanganate Analysis,  ','KMnO'[4])))

c <- POX.smp.summ %>% ggplot(aes(y = CV.conc, x = Weight, color = Grind)) + geom_point(size = 6) + labs(x = "Sample Weight", y = 'CV of Concetration')


d <- POX.smp.summ %>% ggplot(aes(y = CV.pox, x = Weight, color = Grind)) + geom_boxplot(aes(fill = Grind)) +labs(x = 'Weight, g', y = 'CV of POX C ')

grid.arrange(a,b,c,d,ncol = 2 ,top = expression(paste('Pottasium Permanganate Analysis','KMnO'[4])))

e <- POX.smp.summ %>% filter(Weight == '2.5' & Grind == '< 2 mm Grind') %>%
  ggplot(aes(y = POX, x = Total.Carbon)) + geom_point(aes(color = pH), size = 6) +
  labs(x = 'Total Carbon, %', y = expression(paste('POX C, ',' mg kg'^-1))) + scale_color_gradientn(colours = terrain.colors(7))
      
 f <- e + scale_x_continuous(limits = c(0,10)) + scale_color_gradient2(low = "black", mid = "white", high = "green", midpoint = 7)


g <- POX.smp.summ %>% filter(Weight == '2.5' & Grind == '< 2 mm Grind') %>%
  ggplot(aes(y = POX, x = Total.Carbon)) + geom_point(aes(color = Clay), size = 6) +
  labs(x = 'Total Carbon, %', y = expression(paste('POX C, ',' mg kg'^-1)))

h <- g + scale_x_continuous(limits = c(0,10)) + scale_color_gradient2(low = "red", mid = "white", high = "blue", midpoint = 20)

grid.arrange(e,f,g,h,ncol = 2, top = "Values within Standard Analysis \n 2.5g, <2mm Grind")

 
B <- POX.smp.summ %>%
  ggplot(aes(x=POX, y = Weight, color = Grind)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  theme_joy() + scale_x_continuous(limits = c(0,2000))+
  labs(x="POX C",y="Weight", title="POX distribution")

B 

R <- POX.smp.summ %>%
  ggplot(aes(x=POX, y = Weight, color = Grind)) + 
  geom_density_ridges(aes(point_color = Grind, point_fill = Grind, point_shape = Grind), alpha = 0.2, jittered_points = TRUE) +
  scale_point_color_hue(l=40) +
  scale_discrete_manual(aesthetics = "point_shape", values = c(21,22,23,24,25))

R


Rs <- POX.smp.summ %>%
  ggplot(aes(x=POX, color = Weight, y = Grind)) + 
  geom_density_ridges(aes(point_color = Weight, point_fill = Weight, point_shape = Weight), alpha = 0.2, jittered_points = TRUE) +
  scale_point_color_hue(l=40) +
  scale_discrete_manual(aesthetics = "point_shape", values = c(21,22,23,24,25))

Rs

rm(a,b,c,d,e,f,g,h, B)
```

##Measured POX with other variables

```{r info}

POX.info <- POX.smp.summ %>% left_join(POX_info %>% mutate(smp.id = as.factor(smp.id)), by = 'smp.id') %>%
  mutate(Order = toupper(Order)) %>% rename(Master = horz.desgn.master)

i <- POX.info %>% ggplot(aes(y = POX, x = Order)) + geom_boxplot(aes(fill = Weight)) + 
labs(x = 'Soil Order', y = 'POX C ') + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=12))

j <- POX.info %>% ggplot(aes(y = POX, x = Master)) + geom_boxplot(aes(fill = Grind)) + 
labs(x = 'Master Horizon', y = 'POX C ') + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=12))

i
j

rm(POX_info, i, j)
```

###ANOVA
Graphical inspection indicates that both sample size and grind impact POX measurements.  A more formal analysis to detect differences - Treating weights as a class (they are not random); balanced factorial means that the type of ANOVA test done isn't critical.  A general linearized model is to determine the significane of of specific main effects (Weight and Grind) and interactions.

Orders and Master horizon may need to be evaluated - though it may be more likely that removing organics will deal with any issue.  And the relationships are very similar


```{r anova}


a <- aov(POX ~ Grind + Weight + Grind*Weight, data=POX.smp.summ)
summary(a)


at <- glm(POX ~ Grind + Weight + Grind*Weight, data=POX.smp.summ)
summary(at,correlation = TRUE, symbolic.cor = TRUE)
anova(at)




# might add mixed model with Order and Master and variables

#without Histosols


  
a.h <- aov(POX ~ Grind + Weight + Grind*Weight, data=POX.info%>% filter(Order != 'HISTOSOLS' ))
summary(a.h)



```

