---
title: "GRACEnet BD"
author: "Wills"
date: "November 8, 2017"
output: html_document
---

ANalysis of GraceNET data

**GraceNet** - Greenhouse gas Reduction through Agricultural Carbon Enhancement network, Agriculture Research Service project data http://usdaars.maps.arcgis.com/apps/MapSeries/index.html?appid=9415d09247f64ae5bde462a3a9292e6c 


Del Grosso, S. J., J. W. White, G. Wilson, B. Vandenberg, D. L. Karlen, R. F. Follett, J. M. F. Johnson et al. "Introducing the GRACEnet/REAP data contribution, discovery, and retrieval system." Journal of environmental quality 42, no. 4 (2013): 1274-1280.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) #clear previous data

rm(list=ls()) #clear previous data

require("knitr")
knitr::opts_chunk$set(echo = TRUE, comment = "#", warning = FALSE, message = FALSE, error =FALSE, tidy.opts=list(width.cutoff=60),tidy=TRUE)

#opts_knit$set(root.dir = "C:/Users/skye.wills/ownCloud")
workdir <- setwd(getwd()) # personalize the working directory for each user.
opts_knit$set(root.dir = workdir)

list.of.packages <- c("ggplot2", "plyr", "Rcpp", "RColorBrewer", "lattice", "aqp", "stringr", "reshape2", "ggthemes", "tidyverse", "RCurl", "Hmisc", "gridExtra", 'ggmosaic', 'png','rstan','lme4', 'printr', "dplyr","tidyr",  "hexbin",  "ggridge", "ggbeeswarm", "viridis", "gridExtra", "cetcolor", "ggjoy", "readxl", "GGally")

# new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
# if(length(new.packages)) install.packages(new.packages)

library(tidyverse)
library(Rcpp)
library(RColorBrewer)
library(lattice)
library(stringr)
library(ggthemes)
library(RCurl)
library(rstan)
library(lme4)
library(printr)
library(hexbin)
library(ggbeeswarm)
library(viridis)
library(gridExtra)
library(cetcolor)
library(ggjoy)
library(readxl)

```

**GRACEnet** - Agriculture Research Service project data

All available data downloaded for this project.

Acosta-Martinez, V.; Adler, P.; Archer, D.; Arriaga, F.; Baker, J.; Balkcom, K.; Brooks, J.; Cook, K.; Del Grosso, S.; Delgado, J.; Dijkstra, F.; Dungan, R.; Durso, L.; Franzluebbers, A.; Frederick, J.; Gesch, R.; Gollany, H.; Halvorson, A.; Huang, C.; Jin, V.; Johnson, J.; Karlen, D.; Lamb, J.; Lehman, R.; Liebig, M.; Novak, J.; Osborne, S.; Phillips, R.; Roth, G.; Sainju, U.; Sistani, K.; Skinner, R.; Varvel, G.; Venterea, R.; Weyers, S.; Acosta-Martinez, V.; Adeli, A.; Augustine, D.; Barbour, N.; Barsotti, J.; Birrel, S.; Bjorneberg, D.; Blumenthal, D.; Bucholtz, D.; Caesar-TonThat, T.; Cantrell, K.; Chelko, C.; Dell , C.; Derner, J.; Dillard, A.; Ducey, T.; Gross, J.; Hatfield, P.; Jabro, J.; Johnson, H.; Kaplan, N.; Kaspar, T.; Knapp, S.; Kovar, J.; Lauenroth, W.; Lenssen, A.; Leytem, A.; McLaughlin, M.; Milchunas, D.; Miles, D.; Myers, M.; Nichols, K.; Pellack, L.; Polumsky, R.; Reyes-Fox, M.; Sauer, T.; Sigua, G.; Simmons, J.; Sindelar, A.; Smith, D.; Stevens, W.; Stott, D.; Tarkalson, D.; Tomer, M.; Upchurch, D.; Waldron, S.; Watts, D.; Wood, C.; Zobeck, T.. 2017. All; 1980-2017. ver. GPSR_NATRES. Fort Collins, CO: USDA-ARS Natural Resources Database. Project: Natural Resources and Genomics. URL: https://gpsr.ars.usda.gov/natres. File Downloaded: 2017-Oct-20;[02:05:46]. PID: d2e663ee-a9c5-4338-a319-f1db3e9276c7.

****GRACEnet Data Portal****
http://usdaars.maps.arcgis.com/apps/MapSeries/index.html?appid=9415d09247f64ae5bde462a3a9292e6c 


**this code chunk downloads data into the folder you set to BD_data - uncomment to run**

```{r dataDownload}
#set BD_data equal to working directory
#BD_data <- getwd()
#get and combine gracenet data
BD_data <- "D:/Disk 2/BD_eval/"  #shortcut for personal use

# natres.url <- 'https://www.cloudvault.usda.gov/index.php/s/4AY9kGm6TVBD2u6/download'
# ifelse(!file.exists(paste0(BD_data, '/natres.xlsx'),download.file(natres.xlsx, destfile = paste0(BD_data,'/data/natres.csv')), "File already downloaded")

library(readxl)
library(tidyverse)

site <- read_excel(paste0(BD_data, "data/natres.xlsx"), sheet ="FieldSites" )
names(site) <- gsub(" ", "_", names(site)) #fix names so that there are no spaces
names(site)[1] <- "SiteID" #fix names so that Site ID matches Treatments and units

units <- read_excel(paste0(BD_data, "data/natres.xlsx"), sheet ="ExperUnits" ) 
names(units) <- gsub(" ", "_", names(units))
names(units) <- gsub("%", "pct", names(units)) #fix names so that there are no %(special character)
names(units)[1] <- "SiteID" #fix names so that Site ID matches Treatments and sites


trt <- read_excel(paste0(BD_data, "data/natres.xlsx"), sheet ="Treatments" )
names(trt) <- gsub(" ", "_", names(trt))
names(trt)[1] <- "SiteID" #fix names so that Site ID matches sites and units

phys <- read_excel(paste0(BD_data, "data/natres.xlsx"), sheet = "MeasSoilPhys")
names(phys) <- gsub(" ", "_", names(phys))
names(phys) <- gsub("%", "pct", names(phys)) #fix names so that there are no %(special character)
names(phys) <- gsub("/", "per", names(phys)) #fix names so that there are no /(special character)

chem <- read_excel(paste0(BD_data, "data/natres.xlsx"), sheet = "MeasSoilChem")  
names(chem) <- gsub(" ", "_", names(chem))
names(chem) <- gsub("%", "pct", names(chem)) #fix names so that there are no %(special character)
names(chem) <- gsub("/", "per", names(chem)) #fix names so that there are no /(special character)

bio <- read_excel(paste0(BD_data, "data/natres.xlsx"), sheet = "MeasSoilBiol")  
names(bio) <- gsub(" ", "_", names(bio))
names(bio) <- gsub("/", "per", names(bio)) #fix names so that there are no /(special character)


#select relevant fields and join descriptors             
                                                                                                         
grnt <- site %>% select(SiteID, Elev = Elevation_m, MLRA, lat = Latitude_decimal_deg, 
                        lon = Longitude_decimal_deg, Spatial_Description, Native_Veg, 
                        MAP = MAP_mm, MAT = MAT_degC) %>%
             inner_join(units %>% select(SiteID, Treatment_ID, Exp_Unit_ID, Field_ID, Slope_pct,
                                         lat_unit = Latitude, lon_unit = Longitude, 
                                      Soil.taxa = Soil_Classification, size_unit = Exp_Unit_Size_m2) %>%                 inner_join(trt %>%  select(SiteID, Treatment_ID, Treatment_Descriptor,                                                           Rotation_Descriptor, Tillage_Descriptor, Cover_Crop) %>%
         #join to soil measurements
                  full_join(phys %>% select(SiteID, Exp_Unit_ID, Date, Treatment_ID, Upper_cm, Lower_cm,                                             Sand_pct, Silt_pct, Clay_pct, Bulk_Density_gpercm3, 
                                            H2O_Stable_Aggregates_pct) %>%
                    full_join(chem %>% select(SiteID, Exp_Unit_ID, Date, Treatment_ID, Upper_cm, Lower_cm,
                                              Organic_C_gCperkg, TSC_gCperkg, pH) %>%
                        full_join(bio %>% select(SiteID, Exp_Unit_ID, Date, Treatment_ID, Upper_cm, Lower_cm,
                                                 Glucosidase_mgperkgperhr, POM_gCperkg)
                            ))))) %>%
        mutate(Top = Upper_cm, Bottom = Lower_cm,  till = Tillage_Descriptor,Sand = Sand_pct, 
               Silt = Silt_pct, Clay = Clay_pct, BD = Bulk_Density_gpercm3, 
               AggStab = as.numeric(H2O_Stable_Aggregates_pct), SOC = Organic_C_gCperkg/10, 
               bg = as.numeric(Glucosidase_mgperkgperhr), POM = POM_gCperkg)


                                                                                                                                                                                                                            
                                                                                                                                                                                                                            #only keep distinct records (no duplicates)
GRNT <- distinct(grnt)




rm(grnt, bio,chem,phys,site, trt, units)
```

```{r graphs}

# Graphs of BD and SOC

#limit to surface horizons
G <- GRNT %>% filter(!is.na(BD) & !is.na(SOC) & Upper_cm == 0)


G %>% ggplot(aes(x = SOC)) + geom_density()
 
G %>% ggplot(aes(y = BD, x=SOC)) + geom_point() +
  ggtitle('BD and SOC')




#bee swarm
B <- G %>%
  ggplot(aes(x=BD, y = till, color = BD)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_gradientn(colours = cet_pal(10, name = "cyclic_mrybm_35-75_c68_n256"), name = "BD")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Tillage Class",caption="GRACEnet by @wills_skye", title="Surface BD by Tillage Class")

B          

#add SOC measurement for visulaization
B_C <- G %>%
  ggplot(aes(x=BD, y = till, color = SOC)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_viridis()+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Tillage Class",caption="Gracenet by @wills_skye", title="Surface BD by Tillage Class")

B_C

#bee swarm
Bo <- G %>%
  ggplot(aes(x=BD, y = till, color = Lower_cm)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_gradientn(colours = cet_pal(10, name = "cyclic_mrybm_35-75_c68_n256"), name = "Lower Depth (cm)")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Tillage",caption="GRACEnet by @wills_skye", title="SurfaceBulk Density by Tillage")

Bo

Bow <- G %>%
  ggplot(aes(x=BD, y = till, color = BD)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_gradientn(colours = cet_pal(10, name = "cyclic_mrybm_35-75_c68_n256"), name = "BD")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
labs(x="Bulk Density",y="Tillage",caption="GRACEnet by @wills_skye", title="Surface Bulk Density by Tillage")

Bow

#with SOC measurement
BowC <- G%>%
  ggplot(aes(x=BD, y = till, color = SOC)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_viridis(option = "magma")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Tillage",caption="GRACEnet by @wills_skye", title="Surface Bulk Density by Tillage")

BowC

BowT <- G  %>%
  ggplot(aes(x=SOC, y = till, color =Lower_cm)) +  
  geom_quasirandom(alpha=0.5, groupOnX=FALSE) + 
  scale_color_viridis(option = "magma")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Tillage",caption="GRACEnet by @wills_skye", title="Surface SOC by Tillage")

BowT  


```


```{r SOCden}

#create an ordered factor of tillage system to graph with
G <- G %>% 
  filter(!is.na(BD) &!is.na(SOC)) %>%
  mutate(TILL = factor(till, levels = c('Conventional Till', 'Conservation Till', 'Strip Till', 'No Till', 'None')))%>%
  mutate(SOCden = BD*SOC/100, Top = as.numeric(Top), Bottom = as.numeric(Bottom)) 

SOC_BD <- G %>%
ggplot(aes(x=SOC, y = BD, color =TILL)) +  geom_point(size =6, alpha = 0.5) +
theme_joy() + 
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
#labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="Bulk Density of O Horizons \n in Wetlands by Dataset")
 labs(x ='% SOC', y=expression(paste('Bulk Density (g cm'^-3,')')), color = 'Tillage System', title = 'SOC by BD')

SOC_BD


SOCden <- G  %>%
  ggplot(aes(x=SOCden, y = TILL, color =BD)) +  
  geom_quasirandom(alpha=0.5, groupOnX=FALSE) + 
  scale_color_viridis(option = "magma")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Tillage",caption="GRACEnet by @wills_skye", title="SOC Density by Tillage")

SOCden 


Den <- G %>%
 ggplot(aes(x=SOCden, y = Bottom, color =TILL)) +  geom_point(size =6, alpha = 0.5) +
theme_joy() + scale_y_reverse() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
#labs(x="Bulk Density",y="Dataset",caption="@wills_skye")
 labs(x =expression(paste('SOC Density (g cm'^-3,')')), y='Bottom Depth (cm)', color = 'Tillage System', title = 'SOC Density by Sample bottom and tillage system')

Den


Den1 <- G %>%
  mutate(SOCden = BD*SOC/100) %>%    
  filter(!is.na(BD) &!is.na(SOC)) %>%
  ggplot(aes(x=SOCden, y = Clay, color = TILL)) +  geom_point(size =6, alpha = 0.5) +
theme_joy() + 
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
#labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="SOCDensity")
 labs(x =expression(paste('SOC Density (g cm'^-3,')')), y='Clay %', color = 'Tillage Type', title = 'Gracenet Samples')

Den1

```



```{r otherprop}
library(GGally)

SOCBD <- G  %>%
  ggplot(aes(x=SOC, y = BD)) +  geom_point(aes(color = TILL), alpha = 0.5) +facet_wrap(~TILL)
SOCBD

SOC_clay <- G  %>%
  ggplot(aes(x=SOC, y = Clay)) +  geom_point(aes(color = TILL), alpha = 0.5) +facet_wrap(~TILL)
SOC_clay

SOC_bg <- G  %>%
  ggplot(aes(x=SOC, y = bg)) +  geom_point(aes(color = TILL), alpha = 0.5) +facet_wrap(~TILL)
SOC_bg

ggcorr(G %>% select(Sand, Silt, Clay, BD, AggStab, SOC, bg, POM, SOCden))

ggpairs(G%>% select(Clay, BD, AggStab, SOC, bg, POM))

```

##more analysis
I'd like to add something by depth.  There's got to be a depth component to these.........