---
title: "SWS_NWCA"
author: "Wills et al"
date: "January 4, 2018"
output: 
  md_document:
  variant: markdown_github
---

comparing datasets

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls()) #clear previous data

require("knitr")
knitr::opts_chunk$set(echo = TRUE, comment = "#", warning = FALSE, message = FALSE, error =FALSE, tidy.opts=list(width.cutoff=60),tidy=TRUE)

opts_knit$set(root.dir = "C:/Users/skye.wills/ownCloud/AGU_BD-OM")
workdir <- setwd(getwd()) # personalize the working directory for each user.
opts_knit$set(root.dir = workdir)

list.of.packages <- c("ggplot2", "plyr", "Rcpp", "RColorBrewer", "lattice", "aqp", "stringr", "reshape2", "ggthemes", "tidyverse", "RCurl", "Hmisc", "gridExtra", 'ggmosaic', 'ggpmisc', 'png','rstan','lme4', 'printr', "dplyr","tidyr",  "hexbin", "ggbeeswarm", "viridis", "gridExtra", "cetcolor", "ggjoy", "readxl", "sp")

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(tidyverse)
library(Rcpp)
library(RColorBrewer)
library(lattice)
library(stringr)
library(ggthemes)
library(RCurl)
library(rstan)
library(lme4)
library(printr)
library(hexbin)
library(ggbeeswarm)
library(viridis)
library(gridExtra)
library(cetcolor)
library(ggjoy)
library(ggpmisc)
library(readxl)
library(sp)

# when ggjoy is decomisioned replace with  "ggridge", 

```



###Bring NWCA and RaCA datasets with SOC and bulk density from publicly availalbe data.  Add location information which is sensitve and only shared with permission.

**RaCA** - Rapid Carbon Assessment Project.  Information and data available at https://www.nrcs.usda.gov/wps/portal/nrcs/detail/soils/survey/?cid=nrcs142p2_054164

**NWCA** -  National Wetland Condition Assessment. Information and data available at: https://www.nrcs.usda.gov/wps/portal/nrcs/detail/soils/survey/?cid=nrcs142p2_054164
Data for this project was made available as part of the soil property cooperation with Kellogg Soil Survey Laboratory.

**this code chunk downloads data into the folder with this project - uncomment to run**

```{r dataDownload}

raca.url <- 'https://nrcs.box.com/shared/static/nivz52qvsgyalx5zir8sgcu7lpp048pl.csv'
ifelse(!file.exists('RaCA_samples.csv'), download.file(raca.url, destfile = paste0('RaCA_samples.csv') )
       , "File already downloaded")

nwca.url <- 'https://nrcs.box.com/shared/static/u3qahrgycjzenwxykiabh93cnmk4r1e6.csv'
ifelse(!file.exists('nwca.SOCsamples_5-10.csv'), download.file(nwca.url, destfile = paste0('nwca.SOCsamples_5-10.csv')), "File already downloaded")


```


##Read Data into R and Prep
#### Combine dataset into one

* Add location data
* Check field names
* Select relevant fields
* Harmonize datasets
  + Rename to same field names
  + Add SET to denote dataset
  + Add any missing fields (assign values or unknown)
      + LU - to match RaCA W- wetland, C- cropland, P- pastureland, F - Forestland, X- CRP, U added for unknown
  
_Master Horizons (M) set for each dataset in previous data management_
_Texture labels need to be assessed in ensure the same abbreviations are being used_
_wpg for ISCN needs to be converted to volumetric fragment content_


```{r dataPrep}

raca <- read.csv(paste0('RaCA_samples.csv'))

raca.loc <- read.csv('RaCA_data_loc.csv')

raca <- left_join(raca, raca.loc, by =c("rcasiteid" = "RaCA_site"))

raca$CALC_SOC <- ifelse(!is.na(raca$caco3),                      ifelse(raca$caco3>0,raca$c_tot_ncs-(raca$caco3*0.12),raca$c_tot_ncs), raca$c_tot_ncs)
raca$SOC <- ifelse(raca$CALC_SOC<0, 0, raca$CALC_SOC)
r <- raca[!is.na(raca$Measure_BD), c("samp", "upedon", "TOP", "BOT", "Model_desg", "M", "texture", "fragvolc", "SOC", "Measure_BD", "Model_BD", "LU", "Lat", "Lon")]
names(r) <- c("samp_id", "Pedon_ID", "Top", "Bottom", "Model_desg", "M", "Texture",  "CoarseFrag", "SOC", "BD", "RF.BD","LU","Lat", "Lon" )
r$SET <- "RaCA"
r$Clay <- NA



nwca <- read.csv('nwca.SOCsamples_5-10.csv')
nwca.loc <- read.csv('NWCA2011_SITEINFO_081315.csv')

nwca <- left_join(nwca, nwca.loc, by = "UID")

w <- nwca[!is.na(nwca$BD), c( "samp", "Pedon", "top", "bottom", "Model_desg" , "M", "TEXTURE_CLASS", "CLAY", "ROCK_FRAGMENTS", "SOC", "BD", "Model", "AA_CENTER_LAT", "AA_CENTER_LON")]
names(w) <- c("samp_id", "Pedon_ID" ,"Top", "Bottom", "Model_desg", "M", "Texture", "Clay", "CoarseFrag", "SOC", "BD", "RF.BD", "Lat", "Lon")
w$LU <- "W"
w$SET <- "NWCA"
w$Pedon_ID <- as.factor(w$Pedon_ID)

#combine dataset into one - missing values are filled with NA
combo <- rbind(r,w)

#there are a number of really high bulk density numbers - primarily in the ISCN set - need to investigate further
combo <- combo %>% filter(BD<2) 

#knitr::kable(n_M),digits = 1, col.names =names(n_S), caption = "Number of Samples in each Dataset", row.names =F, padding= 2)

```


##Compare Datasets
Use dplyr (with piping) to format and ggplot to create graphs

```{r graphs}
d <- combo %>%
  ggplot(aes(BD)) + geom_density(aes(fill = SET), alpha=.5) + scale_x_continuous(limits = c(0,3)) +
  facet_wrap(~M) + ggtitle('Bulk density by Master Horizon')
d

b <- combo %>%
  #filter(!is.na(BD)) %>%
  ggplot(aes(y = BD, x=SET)) + geom_boxplot(aes(fill=LU)) +
  facet_wrap(~M) + ggtitle('Bulk density by Dataset and Land Use/Cover Class')
b

o <- combo %>%
  filter(!is.na(BD) & M == "O") %>%
  ggplot(aes(y = BD, x=SET)) + geom_boxplot(aes(fill=LU)) +
  ggtitle('For O (organic) horizons only \n Bulk density by Master Horizon and Land Use/Cover Class')
o

s <- combo %>%
  filter(!is.na(BD) & M == "O") %>%
  ggplot(aes(y = BD, x=SOC)) + geom_point(aes(color=LU, shape = SET)) +
  ggtitle('For O (organic) horizons only \n Bulk density and SOC')
s

S <- combo %>%
  filter(!is.na(BD) & M == "O") %>%
  ggplot(aes(y = 1/BD, x=SOC)) + geom_point(aes(color=LU, shape = SET)) +
  ggtitle('For O (organic) horizons only \n Bulk density and SOC')
S

Sw <- combo %>%
  filter(!is.na(BD) & M == "O" & LU == "W") %>%
  ggplot(aes(y = 1/BD, x=SOC)) + geom_point(aes(color=SET, shape = SET)) +
  ggtitle('For O (organic) horizons in Wetlands only \n Bulk density and SOC')
Sw
```


```{r r-plots}

#joy plot with continuous fill
R <- combo %>%
      filter(!is.na(BD)) %>%
      ggplot(aes(x=BD, y = SET, fill = ..x..)) +  
    geom_joy_gradient()  + 
    scale_fill_gradientn(colours = cet_pal(5, name = "fire"), name = "BD")+
    theme_joy() +
    theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
    labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="Ridgeplot of BD by Dataset")

R


#bee swarm
B <- combo %>%
      filter(!is.na(BD)) %>%
  ggplot(aes(x=BD, y = SET, color = BD)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_gradientn(colours = cet_pal(10, name = "cyclic_mrybm_35-75_c68_n256"), name = "BD")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="Beeswarm Plot of BD by Dataset")

B

#add SOC measurement for visulaization
B_C <- combo %>%
      filter(!is.na(BD)&!is.na(SOC) ) %>%
  ggplot(aes(x=BD, y = SET, color = SOC)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_viridis()+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="Beeswarm Plot of BD by Dataset")

B_C

#limit to organic horizons
#bee swarm
Bo <- combo %>%
      filter(!is.na(BD) & M =='O') %>%
  ggplot(aes(x=BD, y = SET, color = BD)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_gradientn(colours = cet_pal(10, name = "cyclic_mrybm_35-75_c68_n256"), name = "BD")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="Bulk Density of Organic Horizons \n by Dataset")

Bo

Bow <- combo %>%
      filter(!is.na(BD) & M =='O' & LU =='W') %>%
  ggplot(aes(x=BD, y = SET, color = BD)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_gradientn(colours = cet_pal(10, name = "cyclic_mrybm_35-75_c68_n256"), name = "BD")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="Bulk Density of O Horizons \n in Wetlands by Dataset")

Bow

#with SOC measurement
BoC <- combo %>%
      filter(!is.na(BD) &!is.na(SOC) & M =='O') %>%
  ggplot(aes(x=BD, y = SET, color = SOC)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_viridis(option = "viridis")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="Bulk Density of O Horizons \n by Dataset")

BoC

BowC <- combo %>%
      filter(!is.na(BD) &!is.na(SOC) & M =='O' & LU =='W') %>%
  ggplot(aes(x=BD, y = SET, color = SOC)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_viridis(option = "viridis")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="Bulk Density of O Horizons \n in Wetlands by Dataset")

BowC

BowCc <- combo %>%
      filter(!is.na(BD) &!is.na(SOC) & M =='O' & LU =='W' & SOC>20) %>%
  ggplot(aes(x=BD, y = SET, color = SOC)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_viridis(option = "viridis")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=14))+
  labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="Bulk Density of O Horizons with SOC > 20% \n in Wetlands by Dataset")

BowCc

BowT <- combo %>%
      filter(!is.na(BD) &!is.na(SOC) & M =='O' & LU =='W') %>%
  ggplot(aes(x=BD, y = SET, color = Top)) +  
  geom_quasirandom(alpha=0.5, groupOnX=FALSE) + 
  scale_color_viridis(option = "magma")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="Bulk Density of O Horizons \n in Wetlands by Dataset")

BowT

BowTc <- combo %>%
      filter(!is.na(BD) &!is.na(SOC) & M =='O' & LU =='W' & SOC > 20) %>%
  ggplot(aes(x=BD, y = SET, color = Top)) +  
  geom_quasirandom(alpha=0.5, groupOnX=FALSE) + 
  scale_color_viridis(option = "magma")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="Bulk Density of O Horizons \n in Wetlands by Dataset")

BowTc

#prep select graphs for export
B <- combo %>%
      filter(!is.na(BD)) %>%
  ggplot(aes(x=BD, y = SET, color = BD)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_gradientn(colours = cet_pal(10, name = "cyclic_mrybm_35-75_c68_n256"), guide = FALSE)+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
  #labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="Beeswarm Plot of BD by Dataset")
 labs(x =expression(paste('Bulk Density (g cm'^-3,')')), y='Dataset')

B

# ggsave('BDswarm_all.png', plot = B, device = "png", path = 'C:/Users/skye.wills/ownCloud/AGU_BD-OMgraphs/poster', scale = 1, width = 6, height = 3, units = "in", dpi = 600, limitsize = TRUE)


BowC <- combo %>%
      filter(!is.na(BD) &!is.na(SOC) & M =='O' & LU =='W') %>%
  ggplot(aes(x=BD, y = SET, color = SOC)) +  
  geom_quasirandom(alpha=0.82, groupOnX=FALSE) + 
  scale_color_viridis(option = "viridis")+
  theme_joy() +
  theme(plot.caption=element_text(hjust=0,size=9),
        axis.title=element_text(size=12),
        plot.title=element_text(size=24))+
#labs(x="Bulk Density",y="Dataset",caption="@wills_skye", title="Bulk Density of O Horizons \n in Wetlands by Dataset")
 labs(x =expression(paste('Bulk Density (g cm'^-3,')')), y='Dataset', color = 'SOC (%)')

BowC




```
##Locations
Summarize by pedons, select all wetlands and plot locations
How do we co-locate?


```{r loc}

#one location per observation/pedon
c <- combo %>%
  filter(LU == "W") %>%
  select(Pedon_ID, Lat, Lon) %>%
  group_by(Pedon_ID) %>%
 summarize_all('min')
  
#create spatial object 

 
```

  