---
title: "Final Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(maps)
library (ggplot2)
library (chron)
library(plyr)
library(scales)
```

## R Markdown

Project Area Location

```{r cars}
map('county', 'ohio') # plot county boundaries for all of Ohio
map('county', 'ohio,madison', add=T, fill = T, col="purple") #add Madison County, Ohio in purple to map
map.text("county", "ohio,Madison", add=T) #add text to madison

#Kokomo <- Sys.glob("C:/workspace/merge71/*.csv") #finds all csv files in folder
#for (Ko in Kokomo) {Kokomo71 <- read.csv(Ko, header = TRUE, sep = ",")} #iterate across all files?
Kokomo71=ldply(list.files(path="C:/workspace/merge71/",pattern="csv",full.names=TRUE),function(filename) {
  dum=read.csv(filename)
  dum$filename=filename
  return(dum)
})

Kokomo71$date <- as.Date (Kokomo71$date, '%m/%d/%Y')

ggplot(data=Kokomo71, aes (date, feet)) +geom_line()+ ylim(10, -5) #plot line

```

**Objectives:**

* Show multiple years of data from one data logger site
* Organize data
* Data Gaps
* Use R instead of excel to create graphs
* Create Plots
  * Totals by months and years
  * Averages by Month and years
  * Show a week of all depths
* Create output report using R Markdown

**Plots:**

```{r pressure, echo=TRUE}
Kokomo71$year <- as.Date(cut(Kokomo71$date,
                              breaks = "year"))
Kokomo71$month <- as.Date(cut(Kokomo71$date,
                              breaks = "month"))
Kokomo71$week <- as.Date(cut(Kokomo71$date,
                             breaks = "week",
                             start.on.monday = FALSE)) # changes weekly break point to Sunday


# graph by year:
ggplot(data = Kokomo71,
       aes(year, feet)) +
  stat_summary(fun.y = "mean", # averages all observations for the year
               geom = "line") + # or "line"
  scale_x_date(date_labels = "%Y", date_breaks = "1 year")

ggplot(data = Kokomo71,
       aes(year, feet)) +
  stat_summary(fun.y = "mean", # averages all observations for the year
               geom = "bar") + # or "line"
  scale_x_date(date_labels = "%Y", date_breaks = "1 year")

# graph by month:
ggplot(data = Kokomo71,
       aes(month, feet)) +
  stat_summary(fun.y = "mean", # averages all observations for the month
               geom = "line") + # or "line"
  scale_x_date(date_labels = "%m/%Y", date_breaks = "3 months")

ggplot(data = Kokomo71,
       aes(month, feet)) +
  stat_summary(fun.y = "mean", # averages all observations for the month
               geom = "bar") + # or "line"
  scale_x_date(date_labels = "%m/%Y", date_breaks = "1 month")

# graph by week:
ggplot(data = Kokomo71,
       aes(week, feet)) +
  stat_summary(fun.y = "mean", # averages all observations for the week
               geom = "line") + # or "line"
  scale_x_date(date_labels = "%d/%m/%Y", date_breaks = "1 week")

Kokomo71$inches <-(Kokomo71$feet*12) #feet x inches
Kokomo71$cm <- (Kokomo71$inches*2.54) #inches x centimeters

# graph by year:
ggplot(data = Kokomo71,
       aes(year, cm)) +
  stat_summary(fun.y = "mean", # averages all observations for the year in centimeters
               geom = "line") + # or "line"
  scale_x_date(date_labels = "%Y", date_breaks = "1 year")

ggplot(data = Kokomo71,
       aes(year, inches)) +
  stat_summary(fun.y = "mean", # averages all observations for the year in inches
               geom = "line") + # or "line"
  scale_x_date(date_labels = "%Y", date_breaks = "1 year")


plot(as.Date(Kokomo71$date,'%d/%m/%Y'),Kokomo71$feet, xlab="Date", ylab= "Feet",type="l", lwd=2, col='purple', main="Kokomo 71")
grid(col="darkgrey")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
