# Lab pedon report
```{r enter soil series name}
# Set soil series
series <- "Genesee"
```

```{r load packages, include=FALSE}
# load libraries
library(aqp)
library(soilDB)
library(reshape)
library(plyr)
library(lattice)
library(maps)
library(xtable)
library(RODBC)
```

```{r fetch and format, load-data, echo=FALSE, warning=FALSE}
# source custom R functions
source(paste0("./genhz_rules/", series, "_rules.R"))

p <- c(0, 0.10, 0.5, 0.90, 1)

# load NASIS data
l <- fetchNASISLabData()
lh <- horizons(l)
lp <- site(l)
f <- fetchNASIS()
h <- horizons(f)
s <- site(f)

names(lh) <- unlist(strsplit(names(lh), "measured"))
lh$cec7Clay <- lh$cec7/lh$claytot

for(i in seq(nrow(lh))){
  if(is.na(lh$hzname[i])) {lh$hzname[i] <-lh$hznameoriginal[i]}
}

lh$hzname[is.na(lh$hzname)] <- "NA"
lh$genhz <- generalize.hz(lh$hzname, ghr[[series]]$n, ghr[[series]]$p)

h$hzname[is.na(h$hzname)] <- "NA"
h$genhz <- generalize.hz(h$hzname, ghr[[series]]$n, ghr[[series]]$p)

naReplace <- function(x){
  l <- list()
  for(i in seq(names(x))){
    if(class(x[,i])=="character") {l[[i]] <- replace(x[,i], is.na(x[,i]), "NA")} else(l[[i]] <-  x[,i])
  }
  l <- data.frame(l, stringsAsFactors=FALSE)
  names(l) <- names(x)
  return(l)
}

h <- naReplace(h)

# Function
sum5n <- function(x) {
  variable <- unique(x$variable)
  precision.vars <- c('phfield', 'ph1to1h2o', 'ph01mcacl2', 'phoxidized', 'ph2osoluble', 'ecec', 'cec7', 'cecsumcations', 'sumbases', 'extracid', 'dbthirdbar', 'dbovendry', 'wthirdbarclod', 'wfifteenbar', 'wretentiondiffws', 'wfifteenbartoclay', 'cec7Clay')
  precision.table <- c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 1, 1, 1, 1, 2)
  v <- na.omit(x$value) # extract column, from long-formatted input data
  precision <- if(variable %in% precision.vars) precision.table[match(variable, precision.vars)] else 0
  n <- length(v)
  ci <- quantile(v, na.rm=TRUE, probs=p) 
  d <- data.frame(min=ci[1], low=ci[2], rv=ci[3], high=ci[4], max=ci[5], n=n, stringsAsFactors=FALSE) # combine into DF
  d$range <- with(d, paste0("(", paste0(round(c(min, low, rv, high, max), precision), collapse=", "), ")", "(", n, ")")) # add 'range' column for pretty-printing
  return(d[7])
}
```

## Plot of all pedons in selected set
Pedons that do not have their Std_Latitude and Std_Longitude columns populated in the NASIS Site table are currently not ploted on the map.
```{r plot of pedons and locations, echo=FALSE}
if(dim(s)[1] != 0) {
  s.sub <- s[complete.cases(s[, c('x_std', 'y_std')]),]
  coordinates(s.sub) <- ~x_std+y_std
  plot(s.sub, pch=16)
  map("county", lwd=0.5, add=T)
  map("state",lwd=2, add=T)
  map.axes()
  } else("no coordinates")
```

## Soil profile plots (depth, color, horizonation, and user pedon id)
```{r Soil plots, echo=FALSE}
plot(l, label="labpeiid")
```

# Summary of NCSS Pedon Lab Data
```{r format site data, echo=FALSE, results='asis'}
# Site information
print(xtable(subset(s, select=c("pedon_id", "taxonname", "tax_subgroup", "part_size_class", "pedon_type", "describer"))), type="html")
```


## Range in characteristics of NCSS Pedon Lab Data
Five number summary (min, 25th, median, 75th, max)(percentiles)
```{r, echo=FALSE, results='asis'}
# Summarize site data
lp.sub <- subset(lp, select=c("noncarbclaywtavg", "claytotwtavg", "le0to100", "wf0175wtavgpsc", "volfractgt2wtavg", "cec7clayratiowtavg"))
lp.lo <- melt(lp.sub, measure.vars=c("noncarbclaywtavg", "claytotwtavg", "le0to100", "wf0175wtavgpsc", "volfractgt2wtavg", "cec7clayratiowtavg"))
lp.5n <- ddply(lp.lo, .(variable), .fun=sum5n)

print(xtable(cast(lp.5n, ~variable, value='range')), type="html")
```


## Box plots of NCSS Pedon Lab Data
Graphical five number summary (outliers, 5th, 25th, median, 75th, 95th, outliers)
```{r bwplot for pedon lab data, echo=FALSE}
if(any(!is.na(lp.lo$value))) {
  bwplot(variable ~ value, data=lp.lo, scales=list(x="free"), xlab="percent")
  } else("no pedon lab data have been populated")         
```


# Summary of NCSS Layer Lab Data

## Horizon designations by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(lh$genhz, lh$hzname)), digits=0, align=rep("c", 2+length(unique(lh$hzname)))), type="html") 
```


## Range in characteristics for generic horizons 
Five number summary (min, 25th, median, 75th, max)(percentiles)
```{r, echo=FALSE}
# Summarize numeric variables by generic horizon
lh.num <- lh[, c(12:(ncol(lh)-4), ncol(lh)-1, ncol(lh))]
lh.num <- Filter(f=function(x) !all(is.na(x)), x=lh.num)
lh.lo <- melt(lh.num, id.vars="genhz", measure.vars=names(lh.num)[c(1:(ncol(lh.num)-1))])
lh.5n <- ddply(lh.lo, .(variable, genhz), .fun=sum5n)
lh.c <- cast(lh.5n, genhz~variable, value='range')

format(lh.c, justify="centre")

nh <- ncol(lh.num)/2.5
```


## Box plots of numeric variables by generic horizon
Graphical five number summary (outliers, 5th, 25th, median, 75th, 95th, outliers)

```{r, echo=FALSE, fig.height=nh}
bwplot(factor(genhz, levels=levels(lh$genhz)[length(levels(lh$genhz)):1]) ~ value|variable, data=rbind(lh.lo), scales=list(x="free"))          
```


## Texture by generic horizon
Contigency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(lh$genhz, lh$lab_texcl)), digits=0, align=rep("c", 1+length(unique(lh$lab_texcl)))), type="html")
```


## Stratified flag by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(lh$genhz, lh$stratextsflag)), digits=0), type="html") 
```


## Depths and thickness of generic horizons
Five number summary (min, 25th, median, 75th, max)(percentiles)

```{r, echo=FALSE, results='asis'}
genhz.thk <- ddply(lh, .(labpeiid, genhz), summarize, thickness=sum(hzdepb-hzdept))
genhz.lo <- melt(genhz.thk, id.vars="genhz", measure.vars="thickness")

lh.lo <- melt(lh, id.vars="genhz", measure.vars=c("hzdept", "hzdepb"))

lp.lo1 <- melt(lp, id.vars="labpeiid", measure.vars=c("psctopdepth", "pscbotdepth"))
lp.thk <- ddply(lp, .(labpeiid), summarize, thickness=sum(pscbotdepth-psctopdepth))
lp.lo2 <- melt(lp.thk, id.vars="labpeiid", measure.vars="thickness")

lh.lo <- rbind(lh.lo, genhz.lo)
lp.lo <- rbind(lp.lo1, lp.lo2)

lh.5n <- ddply(lh.lo, .(variable, genhz), .fun=sum5n)
lh.c <- cast(lh.5n, genhz ~ variable, value='range')
lp.5n <- ddply(lp.lo, .(variable), .fun=sum5n)
print(xtable(lh.c, digits=0, align=rep("c", 1+ncol(lh.c))), type="html")
print(xtable(lp.5n, digits=0, align=rep("c", 1+ncol(lp.5n))), type="html")
```


## Boxplot of generic horizon thicknesses
Graphical five number summary (outliers, 5th, 25th, median, 75th, 95th, outliers)(percentiles)

```{r, echo=FALSE}
bwplot(factor(genhz, levels=levels(lh$genhz)[length(levels(lh$genhz)):1]) ~ value|variable, data=lh.lo, scales=list(x="free"), xlab="cm", horizontal=T)          

bwplot(factor(variable, levels=levels(lp.lo$variable)[length(levels(lp.lo$variable)):1]) ~ value|variable, data=lp.lo, scales=list(x="free"), xlab="cm", horizontal=T)          
```
