
# Map Unit GIS Summary Report

This report summarizes the geographic setting of a list of MUKEY within a file geodatabase. It is intented to be used to compare and contrast map units, and suggest possible Low, RV, and High values for components. 


```{r load-data, echo=FALSE, results='hide', message=FALSE}
mukey <- c('162342') # List of mukey
dir <- "M:/geodata/project_data/"
geodatabase <- "RTSD_Region_11_IND_FY15.gdb"
office <- "11IND"

# Defaults
gdalpath <- "C:/Program Files/QGIS Dufour/bin"
samplePercent <- 10
p <- c(0,0.25,0.5,0.75,1)

pd <- paste0(dir, office, "/")
rd <- paste0(dir, "11REGION/")
shp <- paste("mu", noquote(paste(mukey[c(1, length(mukey))], collapse=".", sep="")), sep="")
cache <- paste0(pd, "cache/", shp, ".shp")
```


```{r LoadPackages, include=F}
library(circular)
library(gdalUtils)
library(sp)
library(raster)
library(rgdal)
library(reshape)
library(plyr)
library(lattice)
library(xtable)
library(maps)

gdal_setInstallation(search_path=gdalpath, rescan=T)
```
```{r map unit summary function, echo=FALSE}
## local functions
mapunit.summary<-function(mukey, samplePercent){
  # Import geodatabase
  if(ifelse(file.exists(cache), "NA", 1) == 1) {
   ogr2ogr(
      src_datasource_name=paste0(pd, geodatabase),
      dst_datasource_name=cache,
      layer="MUPOLYGON",
      where=paste("MUKEY IN (", noquote(paste("'", mukey, "'", collapse=",", sep="")),")", sep=""),
      s_srs=CRS("+init=epsg:5070"),
      t_srs=CRS("+init=epsg:5070"),
      overwrite=T,
      verbose=TRUE
      )
    sapolygon <- readOGR(dsn=paste0(pd, geodatabase), layer="SAPOLYGON", encoding="OpenFileGDB")
    mupolygon <- readOGR(dsn=cache, layer=shp, encoding="ESRI Shapefile")
    mupolygon@data$Acres <- mupolygon@data$Shape_Area * 0.000247105
    mupolygon@data$Circularity <- mupolygon@data$Shape_Leng/(2*sqrt(mupolygon@data$Shape_Area/pi)*pi)
    n <- sum(data.frame(mupolygon)$Shape_Area)/900*samplePercent/100

    # Sample soil map units
    mupolygon.s<-spsample(mupolygon, n=n, "stratified")
    mapunit.df<-cbind(mupolygon.s,over(mupolygon.s,mupolygon))

    # Load grids
    g10.st <- stack(
      paste0(pd, "ned10m_", office, "_slope5.tif"), 
      paste0(pd, "ned10m_", office, "_aspect5.tif")
      )
    g30.st <- stack(
      paste0(pd, "ned30m_", office, ".tif"),
      paste0(pd, "ned30m_", office, "_wetness.tif"),
      paste0(pd, "ned30m_", office, "_mvalleys.tif"),
      paste0(pd, "ned30m_", office, "_z2stream.tif"),
      paste0(pd, "nlcd30m_", office, "_lulc2011.tif")
      )
    g800.st <- stack(
      paste0(rd, "prism800m_11R_ppt_1981_2010_annual_mm.tif"),
      paste0(rd, "prism800m_11R_tmean_1981_2010_annual_C.tif")
      )
    g1000.st <- stack(
      paste0(rd, "rmrs1000m_11R_ffp_1961_1990_annual_days.tif")
      )
  
    names(g10.st) <- c("slope", "aspect")
    names(g30.st) <- c("elev", "wetness", "valley", "relief", "lulc")
    names(g800.st) <- c("ppt", "temp")
    names(g1000.st) <- c("ffp")

  
    # Extract geodata
    g10.e <- extract(g10.st, mupolygon.s)
    g30.e <- extract(g30.st, mupolygon.s)
    g800.e <- extract(g800.st, mupolygon.s)
    g1000.e <- extract(g1000.st, mupolygon.s)
    geodata.df <- data.frame(g10.e, g30.e, g800.e, g1000.e)
  
    # Prep data
    data <- cbind(mapunit.df[, c("MUKEY", "AREASYMBOL", "MUSYM")], geodata.df)
    names(data)[1:3] <- c("MUKEY", "AREASYMBOL", "MUSYM")
    data$aspect <- circular(data$aspect, template="geographic", units="degrees", modulo="2pi")
  
    slope.list<-c(0, 2, 4, 8, 15, 30, 50, 75, 350)
    aspect.list<-c(0, 23, 68, 113, 158, 203, 248, 293, 338, 360) 
    valley.list<-c(0, 0.5, 30)
    lulc.list <- 1:256-1
  
    data$slopeBreaks <- cut(data$slope, breaks=slope.list, right=FALSE)
    data$aspectBreaks <- cut(data$aspect, breaks=aspect.list, right=FALSE)
    data$valleyBreaks <- cut(data$valley, breaks=valley.list, right=FALSE)
    data$lulcBreaks <- cut(data$lulc, breaks=lulc.list, right=FALSE)

    levels(data$slopeBreaks) <- c("0-2","2-4","4-8","8-15","15-30","30-50","50-75","75-350")
    levels(data$aspectBreaks) <- c("N","NE","E","SE","S","SW","W","NW","N")
    levels(data$valleyBreaks) <- c("upland","lowland")
    levels(data$lulcBreaks) <- c('Unclassified','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','Open Water','Perennial Snow/Ice','NA','NA','NA','NA','NA','NA','NA','NA','Developed, Open Space','Developed, Low Intensity','Developed, Medium Intensity','Developed, High Intensity','NA','NA','NA','NA','NA','NA','Barren Land','NA','NA','NA','NA','NA','NA','NA','NA','NA','Deciduous Forest','Evergreen Forest','Mixed Forest','NA','NA','NA','NA','NA','NA','NA','NA','Shrub/Scrub','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','Herbaceuous','NA','NA','NA','NA','NA','NA','NA','NA','NA','Hay/Pasture','Cultivated Crops','NA','NA','NA','NA','NA','NA','NA','Woody Wetlands','NA','NA','NA','NA','Emergent Herbaceuous Wetlands','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA')
  
  data.l <- list(data=data, mupolygon=mupolygon, sapolygon=sapolygon)
  save(data.l, file=paste0(cache, ".Rdata"))
  return(data.l)
  } else load(file=paste0(cache, ".Rdata")) #loading
  return(data.l)
}

conditional.l.rv.h.summary <- function(x) {
  variable <- unique(x$variable)
  v <- na.omit(x$value) # extract column, from long-formatted input data
  precision <- if(variable == 'Circularity') 1 else 0
  ci <- quantile(v, na.rm=TRUE, probs=p) 
  d <- data.frame(min=ci[1], low=ci[2], rv=ci[3], high=ci[4], max=ci[5], stringsAsFactors=FALSE) # combine into DF
  d$range <- with(d, paste("(", paste(round(c(min, low, rv, high, max), precision),collapse=', '), ")", sep="")) # add 'range' column for pretty-printing
  return(d[6])
}

data.l <- mapunit.summary(mukey, samplePercent)
data <- data.l$data
mupolygon <- data.l$mupolygon
sapolygon <- data.l$sapolygon
h <- length(levels(data.l$MUKEY))*3

data$SSA_MUSYM_MUKEY <- as.character(paste(data$AREASYMBOL, data$MUSYM, data$MUKEY))
data2 <- data
data2$MUKEY <- "new_mlra_mapunit"
data2$SSA_MUSYM_MUKEY <- "new_mlra_mapunit"
data <- rbind(data, data2)
```


## Plot of soil map units
```{r plot soil map units, echo=FALSE}
temp <- spTransform(mupolygon, CRS("+init=epsg:4326"))
plot(temp)
map("county", add=T)
map("state",lwd=2,add=T)
map.axes()
```


## Soil polygon metrics
Five number summary (min, 25th, median, 75th, max)(percentiles) and contingency table (counts) 

```{r soil polygon metrics, echo=FALSE, results='asis'}
pol <- data.frame(mupolygon)
pol.m <- melt(pol, id.vars="MUKEY", measure.vars=c("Acres", "Circularity"))
pol.as.m <- melt(pol, id.vars=c("AREASYMBOL", "MUSYM", "MUKEY"), measure.vars=c("Acres"))
pol.as.cs1 <- ddply(pol.as.m, .(AREASYMBOL, MUSYM, MUKEY, variable), .fun=conditional.l.rv.h.summary)
pol.as.cs2 <- ddply(pol.as.m, .(AREASYMBOL, MUSYM, variable), summarize, nPolygons=length(value), nArces=sum(value))[, c(1, 2, 4, 5)]
pol.as.me <- merge(pol.as.cs1, pol.as.cs2, by=c("AREASYMBOL", "MUSYM"))
print(xtable(pol.as.me, digits=0, caption="Summary of MUKEY by AREASYMBOL"), type="html", caption.placement="top")
```

## Polygon Boxplots
Graphical five number summary plus outliers (outliers, 5th, 25th, median, 75th, 95th, outliers)(percentiles)

```{r bwplot of polygon metrics, echo=FALSE}
bwplot(MUKEY ~ value|variable, data=pol.m, scales=list(x="free"))
```

## Percentages
Contingency tables (percent) 

```{r percentage by mukey, echo=FALSE, results='asis'}
## Create descriptive and graphical summary of map unit
print(xtable(prop.table(xtabs(~SSA_MUSYM_MUKEY+slopeBreaks,data=data), margin=1)*100, digits=0, caption="Slope breaks"), type="html", caption.placement="top")
print(xtable(prop.table(xtabs(~SSA_MUSYM_MUKEY+aspectBreaks,data=data), margin=1)*100, digits=0, caption="Aspect breaks"), type="html", caption.placement="top")
print(xtable(prop.table(xtabs(~SSA_MUSYM_MUKEY+valleyBreaks,data=data), margin=1)*100, digits=0, caption="Upland vs. lowland"), type="html", caption.placement="top")
print(xtable(prop.table(xtabs(~SSA_MUSYM_MUKEY+lulcBreaks,data=data), margin=1)*100, digits=0, caption="Landuse and Landcover"), type="html", caption.placement="top")
```

## Percentiles
Five number summary (min, 25th, median, 75th, max)(percentiles)

```{r quantiles by mukey, echo=FALSE, results="asis"}
data.m <- melt(data, id.vars="SSA_MUSYM_MUKEY", measure.vars=c("elev", "slope", "valley", "wetness", "relief", "ppt", "temp", "ffp"))
data.cs <- ddply(data.m, .(variable, SSA_MUSYM_MUKEY), .fun=conditional.l.rv.h.summary)
data.cs.c <- cast(data.cs, SSA_MUSYM_MUKEY ~ variable, value='range')
data.n <- ddply(data, .(SSA_MUSYM_MUKEY), .drop=T, summarize, n=length(MUKEY))
print(xtable(merge(data.cs.c, data.n, by="SSA_MUSYM_MUKEY")), type="html")
aspect.m <- melt(data, id.vars="SSA_MUSYM_MUKEY", measure.vars=c("aspect"))
aspect.m$value <- circular(aspect.m$value, template="geographic", units="degrees", modulo="2pi")
aspect.cs <- ddply(aspect.m, .(variable, SSA_MUSYM_MUKEY), .fun=conditional.l.rv.h.summary)
aspect.cs.c <- cast(aspect.cs, SSA_MUSYM_MUKEY ~ variable, value='range')
print(xtable(merge(aspect.cs.c, data.n, by="SSA_MUSYM_MUKEY")), type="html")
```

## Boxplots of map unit properties
Graphical five number summary plus outliers (outliers, 5th, 25th, median, 75th, 95th, outliers)(percentiles) 

```{r bwplot of map unit properties, echo=FALSE}
### *(min, 5th, 25th, median, 75th, 95th, max)*
temp <- factor(unique(data$SSA_MUSYM_MUKEY))

bwplot(factor(SSA_MUSYM_MUKEY, levels=levels(temp)[length(levels(temp)):1]) ~ value|variable, data=data.m, scales=list(x="free"))          
```

### Variables
Abbreviation| Measure                             | Units
------------|-------------------------------------| -----
elev        | elevation                           | meters
slope       | slope gradient                      | percent
aspect      | slope aspect                        | degrees
valley      | multiresolution valley bottom index | unitless
wetness     | topographic wetness index           | unitless
relief      | height above channel                | meters
ppt         | annual precipitation                | millimeters
temp        | anuual air temperature              | Celsius
ffp         | frost free period                   | days
circularity | shape complexity                    | mupolygon length / mupolygon circumference
