
# Map Unit Spatial Summary Report

This report summarizes the geographic setting of a list of MUKEY within a file geodatabase. It is intented to be used to compare and contrast map units, and suggest possible Low, RV, and High values for components. 

```{r load-data, echo=FALSE, results='hide', message=FALSE}
project <- c('SDJR - MLRA 111A - Genesee silt loam, till substratum, 0 to 2 percent slopes, rarely flooded') # List of project
dir <- "M:/geodata/project_data/"
geodatabase <- "RTSD_Region_11_IND_FY15_copy.gdb"
office <- "11IND"

# Defaults
gdalpath <- "C:/Program Files/QGIS Dufour/bin"
samplePercent <- 15
p <- c(0,0.25,0.5,0.75,1)

pd <- paste0(dir, office, "/")
rd <- paste0(dir, "11REGION/")
shp <- paste0(abbreviate(project, 30))
cache <- paste0(pd, "cache/", shp, ".shp")
```


```{r LoadPackages, include=F}
library(circular)
library(gdalUtils)
library(sp)
library(raster)
library(rgdal)
library(reshape)
library(plyr)
library(lattice)
library(xtable)
library(maps)

gdal_setInstallation(search_path=gdalpath, rescan=T)
```
```{r map unit summary function, echo=FALSE}
## local functions
mapunit.summary<-function(project, samplePercent){
  # Import geodatabase
  if(ifelse(file.exists(paste0(cache, ".Rdata")), FALSE, TRUE)) {
   ogr2ogr(
      src_datasource_name=paste0(pd, geodatabase),
      dst_datasource_name=cache,
      layer="Project_Record",
      where=paste("PROJECT_NAME IN (", noquote(paste("'", project, "'", collapse=",", sep="")),")", sep=""),
      s_srs=CRS("+init=epsg:5070"),
      t_srs=CRS("+init=epsg:5070"),
      overwrite=T,
      simplify=2,
      verbose=TRUE
      )
    sapolygon <- readOGR(dsn=paste0(pd, geodatabase), layer="SAPOLYGON", encoding="OpenFileGDB")
    mupolygon <- readOGR(dsn=cache, layer=shp, encoding="ESRI Shapefile")
    mupolygon@data$Acres <- mupolygon@data$SHAPE_Area * 0.000247105
    mupolygon@data$Circularity <- mupolygon@data$SHAPE_Leng/(2*sqrt(mupolygon@data$SHAPE_Area/pi)*pi)
    n <- sum(data.frame(mupolygon)$SHAPE_Area)/900*samplePercent/100

    # Sample soil map units
    mupolygon.s<-spsample(mupolygon, n=n, "stratified")
    mapunit.df<-cbind(mupolygon.s,over(mupolygon.s,mupolygon))

    # Load grids
    g10.st <- stack(
      paste0(pd, "ned10m_", office, "_slope5.tif"), 
      paste0(pd, "ned10m_", office, "_aspect5.tif")
      )
    g30.st <- stack(
      paste0(pd, "ned30m_", office, ".tif"),
      paste0(pd, "ned30m_", office, "_wetness.tif"),
      paste0(pd, "ned30m_", office, "_mvalleys.tif"),
      paste0(pd, "ned30m_", office, "_z2stream.tif"),
      paste0(pd, "nlcd30m_", office, "_lulc2011.tif")
      )
    g800.st <- stack(
      paste0(rd, "prism800m_11R_ppt_1981_2010_annual_mm.tif"),
      paste0(rd, "prism800m_11R_tmean_1981_2010_annual_C.tif")
      )
    g1000.st <- stack(
      paste0(rd, "rmrs1000m_11R_ffp_1961_1990_annual_days.tif")
      )
  
    names(g10.st) <- c("slope", "aspect")
    names(g30.st) <- c("elev", "wetness", "valley", "relief", "lulc")
    names(g800.st) <- c("ppt", "temp")
    names(g1000.st) <- c("ffp")

  
    # Extract geodata
    g10.e <- extract(g10.st, mupolygon.s)
    g30.e <- extract(g30.st, mupolygon.s)
    g800.e <- extract(g800.st, mupolygon.s)
    g1000.e <- extract(g1000.st, mupolygon.s)
    geodata.df <- data.frame(g10.e, g30.e, g800.e, g1000.e)
  
    # Prep data
    data <- cbind(mapunit.df[, c("AREASYMBOL", "MUSYM")], geodata.df)
    names(data)[1:2] <- c("AREASYMBOL", "MUSYM")
    data$aspect <- circular(data$aspect, template="geographic", units="degrees", modulo="2pi")
  
    slope.list<-c(0, 2, 6, 12, 18, 30, 50, 75, 350)
    aspect.list<-c(0, 23, 68, 113, 158, 203, 248, 293, 338, 360) 
    valley.list<-c(0, 0.5, 30)
    lulc.list <- 1:256-1
  
    data$slopeBreaks <- cut(data$slope, breaks=slope.list, right=FALSE)
    data$aspectBreaks <- cut(data$aspect, breaks=aspect.list, right=FALSE)
    data$valleyBreaks <- cut(data$valley, breaks=valley.list, right=FALSE)
    data$lulcBreaks <- cut(data$lulc, breaks=lulc.list, right=FALSE)

    levels(data$slopeBreaks) <- c("0-2","2-6","6-12","12-18","18-30","30-50","50-75","75-350")
    levels(data$aspectBreaks) <- c("N","NE","E","SE","S","SW","W","NW","N")
    levels(data$valleyBreaks) <- c("upland","lowland")
    levels(data$lulcBreaks) <- c('Unclassified','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','Open Water','Perennial Snow/Ice','NA','NA','NA','NA','NA','NA','NA','NA','Developed, Open Space','Developed, Low Intensity','Developed, Medium Intensity','Developed, High Intensity','NA','NA','NA','NA','NA','NA','Barren Land','NA','NA','NA','NA','NA','NA','NA','NA','NA','Deciduous Forest','Evergreen Forest','Mixed Forest','NA','NA','NA','NA','NA','NA','NA','NA','Shrub/Scrub','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','Herbaceuous','NA','NA','NA','NA','NA','NA','NA','NA','NA','Hay/Pasture','Cultivated Crops','NA','NA','NA','NA','NA','NA','NA','Woody Wetlands','NA','NA','NA','NA','Emergent Herbaceuous Wetlands','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA','NA')
  
  data.l <- list(data=data, mupolygon=mupolygon, sapolygon=sapolygon)
  save(data.l, file=paste0(cache, ".Rdata"))
  return(data.l)
  } else load(file=paste0(cache, ".Rdata")) #loading
  return(data.l)
}

sum5n <- function(x) {
  variable <- unique(x$variable)
  v <- na.omit(x$value) # extract column, from long-formatted input data
  precision <- if(variable == 'Circularity') 1 else 0
  ci <- quantile(v, na.rm=TRUE, probs=p) 
  d <- data.frame(min=ci[1], low=ci[2], rv=ci[3], high=ci[4], max=ci[5], stringsAsFactors=FALSE) # combine into DF
  d$range <- with(d, paste0("(", paste0(round(c(min, low, rv, high, max), precision), collapse=', '), ")")) # add 'range' column for pretty-printing
  return(d[6])
}

data.l <- mapunit.summary(project, samplePercent)
data <- data.l$data
mupolygon <- data.l$mupolygon
sapolygon <- data.l$sapolygon

n <- length(levels(data$AREASYMBOL))
nh <- ifelse(n < 4, 4, 4+(n-4)*.2*4)

data$SSA_MUSYM <- as.character(paste(data$AREASYMBOL, data$MUSYM))
data2 <- data
data2$SSA_MUSYM <- "*new_mlra_mapunit"
data <- rbind(data, data2)
```


### Project Name
```{r project name, echo=FALSE}
project
```


### Variables
Abbreviation| Measure                             | Units                                      | Source
------------|-------------------------------------|------------------------------------------- |-------
elev        | elevation                           | meters                                     | 30-meter USGS National Elevation Dataset (NED)  
slope       | slope gradient                      | percent                                    | 10-meter NED
aspect      | slope aspect                        | degrees                                    | 10-meter NED
valley      | multiresolution valley bottom index | unitless                                   | 30-meter NED
wetness     | topographic wetness index           | unitless                                   | 30-meter NED  
relief      | height above channel                | meters                                     | 30-meter NED
ppt         | annual precipitation                | millimeters                                | 800-meter 30-year normals (1981-2010) from PRISM Climate Dataset
temp        | anuual air temperature              | degrees Celsius                            | 800-meter 30-year normals (1981-2010) from PRISM Climate Dataset
ffp         | frost free period                   | days                                       | 1000-meter 30-year normals (1961-1990) from USFS RMRS
lulc        | land use and land cover             | landcover class (e.g. Woody Wetlands)      | 2011 National Land Cover Dataset (NLCD)


## Map of soil map units
Don't be concerned if the soil map units don't line up with the counties. The county layer being used is an internal R layer that is highly generalized.
```{r plot soil map units, echo=FALSE}
temp <- spTransform(mupolygon, CRS("+init=epsg:4326"))
plot(temp)
map("county", add=T)
map("state",lwd=3,add=T)
map.axes()
```

## Soil polygon metrics
Five number summary (min, 25th, median, 75th, max)(percentiles) and contingency table (counts)(n) 

```{r soil polygon metrics, echo=FALSE, results='asis'}
pol <- data.frame(mupolygon)
pol$SSA_MUSYM <- as.character(paste(pol$AREASYMBOL, pol$MUSYM))
pol2 <- pol
pol2$SSA_MUSYM <- "*new_mlra_mapunit"
pol <- rbind(pol, pol2)

pol.lo1 <- melt(pol, id.vars="SSA_MUSYM", measure.vars=c("Acres", "Circularity"))
pol.lo2 <- melt(pol, id.vars="SSA_MUSYM", measure.vars=c("Acres"))
pol.5n1 <- ddply(pol.lo2, .(SSA_MUSYM, variable), .fun=sum5n)
pol.5n2 <- ddply(pol, .(SSA_MUSYM), summarize, nArces=round(sum(Acres), 0), nPolygons=length(MUSYM))
pol.5n <- merge(pol.5n1, pol.5n2, by=c("SSA_MUSYM"))
print(xtable(pol.5n, digits=0, , align=rep("c", 1+ncol(pol.5n)), caption="Summary of MUSYM by AREASYMBOL"), type="html", caption.placement="top")
```

## Polygon Boxplots
Graphical five number summary plus outliers (outliers, 5th, 25th, median, 75th, 95th, outliers)(percentiles). 
Circularity is an estimate of shape complexity (Hole and Campbell, 1975), computed as a ratio of mupolygon length / mupolygon circumference. The shape complexity of a perfect circle would equal 1.

```{r bwplot of polygon metrics, echo=FALSE}
bwplot(SSA_MUSYM ~ value|variable, data=pol.lo1, scales=list(x="free"))
```

## Contingency tables (percent) 

```{r percentage by musym, echo=FALSE, results='asis'}
## Create descriptive and graphical summary of map unit
print(xtable(prop.table(xtabs(~SSA_MUSYM+slopeBreaks,data=data), margin=1)*100, align=rep("c", 1+length(levels(data$slopeBreaks))), digits=0, caption="Slope breaks"), type="html", caption.placement="top")
print(xtable(prop.table(xtabs(~SSA_MUSYM+aspectBreaks,data=data), margin=1)*100, align=rep("c", 1+length(levels(data$aspectBreaks))), digits=0, caption="Aspect breaks"), type="html", caption.placement="top")
print(xtable(prop.table(xtabs(~SSA_MUSYM+valleyBreaks,data=data), margin=1)*100, align=rep("c", 1+length(levels(data$valleyBreaks))), digits=0, caption="Upland vs. lowland"), type="html", caption.placement="top")
print(xtable(addmargins(prop.table(xtabs(~SSA_MUSYM+lulcBreaks,data=data), margin=1)*100), digits=0, align=rep("c", 2+length(levels(data$lulcBreaks))), caption="Landuse and Landcover"), type="html", caption.placement="top")
```

## Quantile breaks
Five number summary (min, 25th, median, 75th, max)(percentiles) and number of random samples (n)

```{r quantiles by musym, echo=FALSE, results="asis"}
data.lo <- melt(data, id.vars="SSA_MUSYM", measure.vars=c("elev", "slope", "valley", "wetness", "relief", "ppt", "temp", "ffp"))
data.5n <- ddply(data.lo, .(variable, SSA_MUSYM), .fun=sum5n)
data.c <- cast(data.5n, SSA_MUSYM ~ variable, value='range')
data.n <- ddply(data, .(SSA_MUSYM), .drop=T, summarize, n=length(SSA_MUSYM))
print(xtable(merge(data.c, data.n, by="SSA_MUSYM"), align=rep("c", 2+ncol(data.c))), type="html")

aspect.lo <- melt(data, id.vars="SSA_MUSYM", measure.vars=c("aspect"))
aspect.lo$value <- circular(aspect.lo$value, template="geographic", units="degrees", modulo="2pi")
aspect.5n <- ddply(aspect.lo, .(variable, SSA_MUSYM), .fun=sum5n)
aspect.c <- cast(aspect.5n, SSA_MUSYM ~ variable, value='range')
print(xtable(merge(aspect.c, data.n, by="SSA_MUSYM"), align=rep("c", 2+ncol(aspect.c))), type="html")
```

## Boxplots of map unit properties
Graphical five number summary plus outliers (outliers, 5th, 25th, median, 75th, 95th, outliers)(percentiles) 

```{r bwplot of map unit properties, echo=FALSE, fig.height=nh}
### *(min, 5th, 25th, median, 75th, 95th, max)*
temp <- factor(unique(data$SSA_MUSYM))

bwplot(factor(SSA_MUSYM, levels=levels(temp)[length(levels(temp)):1]) ~ value|variable, data=data.lo, scales=list(x="free"))          
```
