# Pedon report
```{r enter soil series name}
# Set soil series
series <- "Genesee"
```


```{r load packages, include=FALSE}
# load libraries
library(aqp)
library(soilDB)
library(reshape)
library(plyr)
library(lattice)
library(maps)
library(xtable)
library(RODBC)
library(circular)
library(RColorBrewer)
```


```{r fetch and format, load-data, echo=FALSE, warning=FALSE}
# source custom R functions
source(paste0("./genhz_rules/", series, "_rules.R"))

p <- c(0, 0.5, 1)

# load NASIS data
pedons <- fetchNASIS(nullFragsAreZero = TRUE)

h <- horizons(pedons)
nai <- is.na(h$genhz)
h[nai, "genhz"] <- as.character(sapply(h[nai, "hzname"], function(x) generalize.hz(x, ghr[[series]]$n, ghr[[series]]$p)))
horizons(pedons) <- h
pedons$genhz[is.na(pedons$genhz)] <- "NA"
pedons$genhz <- factor(pedons$genhz, levels = ghr[[series]]$n)                                                                                                       

h <- horizons(pedons)
s <- site(pedons)
d <- diagnostic_hz(pedons)

h$effervescence <- as.character(h$effervescence)

na_replace <- function(x){
  if(class(x) == "character") {x <- replace(x, is.na(x), "missing")} 
  else (x <-  x)
}

h <- data.frame(lapply(h, na_replace), stringsAsFactors = FALSE)
s <- data.frame(lapply(s, na_replace), stringsAsFactors = FALSE)

l <- length(unique(h$peiid))

# Function
sum5n <- function(x) {
  variable <- unique(x$variable)
  precision.vars <- c('phfield', 'ph1to1h2o', 'ph01mcacl2', 'phoxidized', 'ph2osoluble', 'ecec', 'cec7', 'cecsumcations', 'sumbases', 'extracid', 'dbthirdbar', 'dbovendry', 'wthirdbarclod', 'wfifteenbar', 'wretentiondiffws', 'wfifteenbartoclay', 'cec7Clay')
  precision.table <- c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 1, 1, 1, 1, 2)
  v <- na.omit(x$value) # extract column, from long-formatted input data
  precision <- if(variable %in% precision.vars) precision.table[match(variable, precision.vars)] else 0
  n <- length(v)
  ci <- quantile(v, na.rm=TRUE, probs=p) 
  d <- data.frame(rbind(ci), n = n, stringsAsFactors=FALSE) # combine into DF
  d$range <- with(d, paste0("(", paste0(round(d[-ncol(d)], precision), collapse=", "), ")", "(", n, ")")) # add 'range' column for pretty-printing
  return(d["range"])
}
```


## Map of all pedons in selected set
Pedons that do not have their Std_Latitude and Std_Longitude columns populated in the NASIS Site table are currently not ploted on the map.

```{r Map of pedons and locations, echo=FALSE}
if(dim(s)[1] != 0) {
  s.sub <- s[complete.cases(s[, c('x_std', 'y_std')]),]
  coordinates(s.sub) <- ~x_std+y_std
  plot(s.sub, pch=16)
  map("county", lwd=0.5, add=T)
  map("state",lwd=2,add=T)
  map.axes()
} else("no coordinates")
```


## Soil profile plots (depth, color by genhz, horizonation, and user pedon id)
```{r Soil plots, echo=FALSE}
cols <- brewer.pal(n=length(levels(pedons$genhz)), name="Paired")
# assign a color to each generalized horizon label
hz.names <- levels(pedons$genhz)
pedons$genhz.soil_color <- cols[match(pedons$genhz, hz.names)]
# plot generalized horizons via color and add a legend
plot(pedons, name='hzname', color='genhz.soil_color', label='pedon_id')
```


# Summary of Sites

```{r format site data, results='asis', echo=FALSE}
# Site information
print(xtable(subset(s, select=c("pedon_id", "taxonname", "tax_subgroup", "part_size_class", "pedon_type", "describer"))), type="html")
```

## Range in characteristics of geographic setting
Five number summary (min, 25th, median, 75th, max)(percentiles) and (n)(total number of observations)

```{r, echo=FALSE, results='asis'}
morf <- subset(s, select = c("elev_field", "slope_field"))
morf.lo <- melt(morf, measure.vars= c("elev_field", "slope_field"))
morf.5n <- ddply(morf.lo, .(variable), .fun=sum5n)
morf.wi <- cast(morf.5n, ~variable, value = 'range')
print(xtable(morf.wi, caption="Elevation and slope gradient"), type="html", caption.placement="top")


if(!all(is.na(s$aspect_field))) {
  aspect <- subset(s, select=c("aspect_field"))
  aspect.lo <- melt(aspect, measure.vars=c("aspect_field"))
  aspect.lo$value <- circular(aspect.lo$value, template="geographic", units="degrees", modulo="2pi")
  aspect.5n <- ddply(aspect.lo, .(variable), .fun=sum5n)
  aspect.wi <- cast(aspect.5n, ~variable, value='range')
  print(xtable(aspect.wi, caption="Slope aspect"), type="html", caption.placement="top")
} else("no aspect values have been populated")         


if(!all(is.na(s[, c("pmorigin", "pmkind", "landform.string")]))) {
  pm.lf <- addmargins(table(factor(paste0(s$pmorigin, " ", s$pmkind)), factor(s$landform.string)))
  print(xtable(pm.lf, digits=0, align=rep("c", 1+ncol(pm.lf)), caption="Parent material vs landform"), type="html", caption.placement="top")} else("no values have been populated for parent material or landform")


if(!all(is.na(s[, c("shapedown", "shapeacross")]))) {
  s$shapedown <- factor(s$shapedown, levels=c("Convex", "Linear", "Concave", "missing"))
  s$shapeacross <- factor(s$shapeacross, levels=c("Convex", "Linear", "Concave", "missing"))
  
  print(xtable(addmargins(table(s$shapedown, s$shapeacross)), digits=0, align=rep("c", 6), caption="Down slope (y-axis) vs across slope (x-axis)"), type="html", caption.placement="top")
} else("no slope shape values have been populated")

if(!all(is.na(s[, c("hillslope_pos", "drainagecl")]))) {
  s$hillslope_pos <- factor(s$hillslope_pos, levels=c("Summit", "Shoulder", "Backslope", "Footslope", "Toeslope", "missing"))
  s$drainagecl <- factor(s$drainagecl, levels= c("Excessivelly drained", "Well drained", "Somewhat excessively well drained", "Moderately well drained", "Somewhat poorly drained", "Poorly drained", "Very poorly drained", "Subaqueous drainage", "missing"))
  
  print(xtable(addmargins(table(s$drainagecl, s$hillslope_pos)), digits=0, align=rep("c", 8), caption="Drainage class vs hillslope position"), type="html", caption.placement="top")
} else("no hillslope position or drainage class values have been populated")
```


## Range in characteristics of surface rock fragments
Five number summary (min, 25th, median, 75th, max)(percentiles) and (n)(total number of observations)

```{r, echo=FALSE, results='asis'}
# Summarize surface rock fragments
srf <- subset(s, select=c("surface_fgravel", "surface_gravel", "surface_cobbles", "surface_stones", "surface_boulders", "surface_flagstones", "surface_channers"))
srf$total_srf <- srf$surface_gravel+srf$surface_cobbles+srf$surface_stones+srf$surface_boulders+srf$surface_flagstones+srf$surface_channers
srf$surface_gravel <- srf$surface_gravel-srf$surface_fgravel
srf.lo <- melt(srf, measure.vars=c("total_srf", "surface_fgravel", "surface_gravel", "surface_cobbles", "surface_stones", "surface_boulders", "surface_channers", "surface_flagstones"))
srf.5n <- ddply(srf.lo, .(variable), .fun=sum5n)
srf.wi <- cast(srf.5n, ~variable, value='range')

print(xtable(srf.wi), type="html")
```


## Box plots of surface rock fragments
Graphical five number summary plus outliers (outliers, 5th, 25th, median, 75th, 95th, outliers)

```{r bwplot for rf, echo=FALSE}
if(sum(srf$total_srf) != 0) {
  bwplot(variable~value, data=srf.lo, scales=list(x="free"), xlab="percent") 
  } else("no surface rock fragments have been populated")         
```


# Summary of Pedons
## Horizon designations by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$hzname)), digits = 0), type="html") 
```


## Range in characteristics for generic horizons 
Five number summary (min, 25th, median, 75th, max)(percentiles) and (n)(total number of observations)

```{r, echo=FALSE, results='asis'}
# Summarize numeric variables by generic horizon
names(h)[15:16] <- c("fragvoltotc", "fragvoltot")
var <- c("fine_gravel", "gravel", "cobbles", "stones", "boulders", "paragravel", "paracobbles", "channers", "flagstones")
h[, var] <- lapply(h[var], as.integer)
h$gravel <- h$gravel-h$fine_gravel

h.lo1 <- melt(h, id.vars="genhz", measure.vars = c('clay', 'sand', 'fine_gravel', 'gravel', 'cobbles', 'stones', 'fragvoltot'))
h.lo2 <- melt(h, id.vars="genhz", measure.vars = c('phfield', 'd_value', 'd_chroma', 'm_value', 'm_chroma'))

h.5n1 <- ddply(h.lo1, .(variable, genhz), .fun = sum5n)
h.5n2 <- ddply(h.lo2, .(variable, genhz), .fun = sum5n)

h.wi1 <- cast(h.5n1, genhz ~ variable, value = 'range')
h.wi2 <- cast(h.5n2, genhz ~ variable, value = 'range')

print(xtable(h.wi1, align=rep("c", 1+ncol(h.wi1))), type="html", include.rownames=FALSE)
print(xtable(h.wi2, align=rep("c", 1+ncol(h.wi2))), type="html", include.rownames=FALSE)

n <- length(unique(h$genhz))
nh <- ifelse(n < 4, 4, 4+(n-4)*.15*3)
```


## Box plots of numeric variables by generic horizon
Graphical five number summary plus outliers (outliers, 5th, 25th, median, 75th, 95th, outliers)

```{r, echo=FALSE, fig.height=nh}
bwplot(factor(genhz, levels = levels(h$genhz)[length(levels(h$genhz)):1]) ~ value|variable, data = rbind(h.lo1, h.lo2), scales = list(x="free"))          
```


## Texture by generic horizon
Contigency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$texture_class)), digits=0, align = rep("c", 2+length(unique(h$texture_class)))), type = "html")
```


## Tex Mod & Class by generic horizon
Contigency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$texture)), digits=0, align=rep("c", 2+length(unique(h$texture)))), type = "html")
```


## Dry hue by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$d_hue)), digits=0, align=rep("c", 2+length(unique(h$d_hue)))), type="html")
```


## Moist hue by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$m_hue)), digits=0, align = rep("c", 2+length(unique(h$m_hue)))), type = "html")
```


## Effervescence by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
h$effervescence[h$effervescence == "very slight"] <- "very.slight"
eff <- c(violent = "violent", strong = "strong", slight = "slight", very.slight = "very slight", none = "none", missing = "missing")
h$effervescence <- eff[h$effervescence]
h$effervescence <- factor(h$effervescence, levels = eff)

print(xtable(addmargins(table(h$genhz, h$effervescence)), digits=0, align=rep("c", 2+length(levels(h$effervescence)))), type="html")
```


## Depths and thickness of generic horizons
Five number summary (min, 25th, median, 75th, max)(percentiles) and (n)(total number of observations)

```{r, echo=FALSE, results='asis'}
genhz.thk <- ddply(h, .(phiid, genhz), summarize, thickness=sum(hzdepb-hzdept))
genhz.lo <- melt(h, id.vars="genhz", measure.vars = c('hzdept', 'hzdepb'))
genhz.lo2 <- melt(genhz.thk, id.vars = "genhz", measure.vars = 'thickness')
genhz.lo <- rbind(genhz.lo, genhz.lo2)
genhz.5n <- ddply(genhz.lo, .(variable, genhz), .fun = sum5n)
print(xtable(cast(genhz.5n, genhz ~ variable, value = 'range'), digits = 0, align = rep("c", 2 + ncol(genhz.5n))), type = "html")
```


## Boxplot of generic horizon thicknesses
Graphical five number summary plus outliers (outliers, 5th, 25th, median, 75th, 95th, outliers)(percentiles)

```{r, echo=FALSE}
bwplot(factor(genhz, levels=levels(h$genhz)[length(levels(h$genhz)):1]) ~ value|variable, data = genhz.lo, scales = list(x = "free"), xlab = "cm", horizontal = T)          
```


## Depths and thickness of diagnostic horizons and features
Five number summary (min, 25th, median, 75th, max)(percentiles) and (n)(total number of observations)

```{r, echo=FALSE, results='asis'}
if(dim(d)[1] == 0)
{d <- data.frame(peiid = h$peiid)
d[ c("diag_kind", "featdept", "featdepb")] <- list(diag_kind = as.character("missing"), featdept=as.integer(NA), featdepb = as.integer(NA))}

diag.thk <- ddply(d, .(peiid, diag_kind), summarize, thickness = sum(featdepb - featdept))
diag.lo <- melt(d, id.vars = "diag_kind", measure.vars = c('featdept', 'featdepb'))
diag.lo2 <- melt(diag.thk, id.vars = "diag_kind", measure.vars = 'thickness')

names(s)[47:48] <- c("featdept", "featdepb")
pscs.thk <- ddply(s, .(peiid), summarize, thickness = sum(featdepb - featdept))
pscs.thk <- data.frame(peiid = pscs.thk$peiid, variable = "thickness", value = pscs.thk$thickness)
pscs.lo <- melt(s, id.vars = "peiid", measure.vars = c('featdept', 'featdepb'))
pscs.lo <- rbind(pscs.lo, pscs.thk)
pscs.lo <- data.frame(diag_kind = "particle size control section", variable = pscs.lo$variable, value = pscs.lo$value)

diag.lo <- rbind(diag.lo, diag.lo2, pscs.lo)
diag.5n <- ddply(diag.lo, .(variable, diag_kind), .fun = sum5n)
diag.wi <- cast(diag.5n, diag_kind ~ variable, value = 'range')

print(xtable(diag.wi, digits = 0, align=rep("c", 1+ncol(diag.wi))), type = "html")
```


## Boxplot of diagnostic horizons and features depth
Graphical five number summary plus outliers (outliers, 5th, 25th, median, 75th, 95th, outliers)(percentiles)

```{r, echo=FALSE}
bwplot(diag_kind ~ value|variable, data = diag.lo, scales = list(x = "free"), xlab = "cm")          
```
