# Pedon report
```{r enter soil series name}
# Set soil series
series <- "Rossmoyne"
```

```{r load packages, include=FALSE}
# load libraries
library(aqp)
library(soilDB)
library(reshape)
library(plyr)
library(lattice)
library(maps)
library(xtable)
```

```{r fetch and format, load-data, echo=FALSE, warning=FALSE}
# source custom R functions
source(paste0("./genhz_rules/", series, "_rules.R"))

p <- c(0, 0.25, 0.5, 0.75, 1)

# load NASIS data
f <- fetchNASIS()

h <- horizons(f)
s <- site(f)

h$hzname[is.na(h$hzname)] <- "NA"
h$genhz <- generalize.hz(h$hzname, ghr[[series]]$n, ghr[[series]]$p)

# Functions below need work, intended to preserve pre-existing genhz from comp_layer_id
# h$genhz <- ifelse(is.na(h$genhz), as.character(generalize.hz(h$hzname, ghr[[series]]$n, ghr[[series]]$p)), h$genhz)
# h$hzdepm <- h$hzdepb-h$hzdept
# ds <- ddply(h, .(genhz), summarize, hzdepm = median(hzdepm))
# ds[order(ds$hzdepm), "genhz"]
# h$genhz <- as.factor(h$genhz, levels=c(ghr[[series]]$n, "not-used"))

d_rgb <- horizons(f)[, c("d_r", "d_g", "d_b")]
d_munsell <- rgb2munsell(d_rgb)
m_rgb <- h[, c("m_r", "m_g", "m_b")]
m_munsell <- rgb2munsell(m_rgb)

h$d_hue <- as.character(d_munsell$hue)
h$d_value <- as.integer(d_munsell$value)
h$d_chroma <- as.integer(d_munsell$chroma)
h$m_hue <- as.character(m_munsell$hue)
h$m_value <- as.integer(m_munsell$value)
h$m_chroma <- as.integer(d_munsell$chroma)

names(h)[15] <- "fragvoltotc"
names(h)[16] <- "fragvoltot"

h[, c("fine_gravel", "gravel", "cobbles", "stones", "boulders", "paragravel", "paracobbles", "channers", "flagstones")] <- lapply(h[, c("fine_gravel", "gravel", "cobbles", "stones", "boulders", "paragravel", "paracobbles", "channers", "flagstones")], as.integer)

h$effervescence <- as.character(h$effervescence)

naReplace <- function(h){
  l <- list()
  for(i in seq(names(h))){
    if(class(h[,i])=="character") {l[[i]] <- replace(h[,i], is.na(h[,i]), "NA")} else(l[[i]] <-  h[,i])
  }
  l <- data.frame(l, stringsAsFactors=FALSE)
  names(l) <- names(h)
  return(l)
}

h <- naReplace(h)

d <- diagnostic_hz(f)

if(dim(d)[1] == 0)
{d <- data.frame(peiid=h$peiid)
d[ c("diag_kind", "featdept", "featdepb")] <- list(diag_kind=as.character("NA"), featdept=as.integer(NA), featdepb=as.integer(NA))}


l <- length(unique(h$peiid))

# Function
conditional.l.rv.h.summary <- function(x) {
  variable <- unique(x$variable)
  v <- na.omit(x$value) # extract column, from long-formatted input data
  precision <- if(variable == 'phfield') 1 else 0
  ci <- quantile(v, na.rm=TRUE, probs=p) 
  d <- data.frame(min=ci[1], low=ci[2], rv=ci[3], high=ci[4], max=ci[5], stringsAsFactors=FALSE) # combine into DF
  d$range <- with(d, paste("(", paste(round(c(min, low, rv, high, max), precision), collapse=', '), ")", sep="")) # add 'range' column for pretty-printing
  return(d[6])
}
```
## Map of all pedons in selected set
Pedons that do not have their Std_Latitude and Std_Longitude columns populated in the NASIS Site table are currently not ploted on the map.

```{r Map of pedons and locations, echo=FALSE}

if(dim(s)[1] != 0) {
  s.sub <- s[complete.cases(s[, c('x_std', 'y_std')]),]
  coordinates(s.sub) <- ~x_std+y_std
  plot(s.sub, pch=16)
  map("county", lwd=0.5, add=T)
  map("state",lwd=1,add=T)
  map.axes()
  } else("no coordinates")
```

## Soil profile plots (depth, color, horizonation, and user pedon id)
```{r Soil plots, echo=FALSE}
plot(f, label="pedon_id")
```


# Summary of Sites
```{r format site data, results='asis', echo=FALSE}
# Site information
print(xtable(subset(s, select=c("pedon_id", "taxonname", "tax_subgroup", "part_size_class", "pedon_type", "describer"))), type="html")

# Summarize surface rock fragments
s.rf <- subset(s, select=c("surface_fgravel", "surface_gravel", "surface_cobbles", "surface_stones", "surface_boulders", "surface_flagstones", "surface_channers"))
s.rf$total_srf <- s.rf$surface_gravel+s.rf$surface_cobbles+s.rf$surface_stones+s.rf$surface_boulders+s.rf$surface_flagstones+s.rf$surface_channers
s.rf$surface_gravel <- s.rf$surface_gravel-s.rf$surface_fgravel
s.m <- melt(s.rf, measure.vars=c("total_srf", "surface_fgravel", "surface_gravel", "surface_cobbles", "surface_stones", "surface_boulders", "surface_channers", "surface_flagstones"))
s.summary <- ddply(s.m, .(variable), .fun=conditional.l.rv.h.summary)
```


## Box plots of surface rock fragments
Graphical five number summary plus outliers (outliers, 5th, 25th, median, 75th, 95th, outliers)

```{r bwplot for rf, echo=FALSE}
bwplot(variable~value, data=s.m, scales=list(x="free"), xlab="percent")          
```


## Range in characteristics of surface rock fragments
Five number summary (min, 25th, median, 75th, max)(percentiles)

```{r, echo=FALSE, results='asis'}
print(xtable(cast(s.summary, ~variable, value='range')), type="html")
```

# Summary of Pedons
```{r, echo=FALSE, results='asis'}
# Summarize numeric variables by generic horizon
h$gravel <- h$gravel-h$fine_gravel

hp.m <- melt(h, id.vars="genhz", measure.vars=c('clay', 'sand', 'fine_gravel', 'gravel', 'cobbles', 'stones', 'fragvoltot'))
hp.cs <- ddply(hp.m, .(variable, genhz), .fun=conditional.l.rv.h.summary)

hc.m <- melt(h, id.vars="genhz", measure.vars=c('phfield', 'd_value', 'd_chroma', 'm_value', 'm_chroma'))
hc.cs <- ddply(hc.m, .(variable, genhz), .fun=conditional.l.rv.h.summary)

thk <- ddply(h, .(peiid, genhz), summarize, thickness=sum(hzdepb-hzdept))
thk.m <- melt(thk, id.vars="genhz", measure.vars="thickness")
thk.cs <- ddply(thk.m, .(variable, genhz), .fun=conditional.l.rv.h.summary)

hp.c <- cast(hp.cs, genhz ~ variable, value='range')
hc.c <- cast(hc.cs, genhz ~ variable, value='range')
```


## Horizon designations by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$hzname)), digits=0), type="html", width=10) 
```


## Range in characteristics for generic horizons 
Five number summary (min, 25th, median, 75th, max)(percentiles)

```{r, echo=FALSE, results='asis'}
print(xtable(hp.c), type="html")
print(xtable(hc.c), type="html")
```

## Box plots of numeric variables by generic horizon
Graphical five number summary plus outliers (outliers, 5th, 25th, median, 75th, 95th, outliers)

```{r, echo=FALSE}
bwplot(factor(genhz, levels=levels(h$genhz)[length(levels(h$genhz)):1]) ~ value|variable, data=rbind(hp.m, hc.m, thk.m), scales=list(x="free"))          
```

## Texture by generic horizon
Contigency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$texture_class)), digits=0), type="html")
```

## Tex Mod & Class by generic horizon
Contigency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$texture)), digits=0), type="html")
```

## Dry hue by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$d_hue)), digits=0), type="html")
```

## Moist hue by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$m_hue)), digits=0), type="html")
```

## Effervescence by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$effervescence)), digits=0), type="html")
```


## Depths and thickness of generic horizons
Five number summary (min, 25th, median, 75th, max)(percentiles)

```{r, echo=FALSE, results='asis'}
genhz.thk <- ddply(h, .(phiid, genhz), summarize, thickness=sum(hzdepb-hzdept))
genhz.m <- melt(h, id.vars="genhz", measure.vars=c('hzdept', 'hzdepb'))
genhz.thk.m <- melt(genhz.thk, id.vars="genhz", measure.vars='thickness')
genhz.m <- rbind(genhz.m, genhz.thk.m)
d.cs <- ddply(genhz.m, .(variable, genhz), .fun=conditional.l.rv.h.summary)
print(xtable(cast(d.cs, genhz ~ variable, value='range'), digits=0), type="html")
```


## Boxplot of generic horizon thicknesses
Graphical five number summary plus outliers (outliers, 5th, 25th, median, 75th, 95th, outliers)(percentiles)

```{r, echo=FALSE}
bwplot(factor(genhz, levels=levels(h$genhz)[length(levels(h$genhz)):1]) ~ value|variable, data=genhz.m, scales=list(x="free"), xlab="cm", horizontal=T)          
```

## Depths and thickness of diagnostic horizons
Five number summary (min, 25th, median, 75th, max)(percentiles)

```{r, echo=FALSE, results='asis'}
diag.thk <- ddply(d, .(peiid, diag_kind), summarize, thickness=sum(featdepb-featdept))
diag.m <- melt(d, id.vars="diag_kind", measure.vars=c('featdept', 'featdepb'))
diag.thk.m <- melt(diag.thk, id.vars="diag_kind", measure.vars='thickness')
diag.m <- rbind(diag.m, diag.thk.m)
d.cs <- ddply(diag.m, .(variable, diag_kind), .fun=conditional.l.rv.h.summary)
print(xtable(cast(d.cs, diag_kind ~ variable, value='range'), digits=0), type="html")
```

## Boxplot of diagnostic horizon thicknesses
Graphical five number summary plus outliers (outliers, 5th, 25th, median, 75th, 95th, outliers)(percentiles)

```{r, echo=FALSE}
bwplot(diag_kind ~ value|variable, data=diag.m, scales=list(x="free"), xlab="cm")          
```

