# Pedon report
```{r enter soil series name}
# Set soil series
series <- "Genesee"
```

```{r load packages, include=FALSE}
# load libraries
library(aqp)
library(soilDB)
library(reshape)
library(plyr)
library(lattice)
library(maps)
library(xtable)
```

```{r fetch and format, load-data, echo=FALSE, warning=FALSE}
# source custom R functions
source("genhz_rules.R")

p <- c(0, 0.10, 0.5, 0.90, 1)

# load NASIS data
f <- fetchNASIS()
f <- subsetProfiles(f, s="taxonname == series")

h <- horizons(f)
s <- site(f)

h$hzname[is.na(h$hzname)] <- "NA"
h$genhz <- generalize.hz(h$hzname, ghr$Genesee$n, ghr$Genesee$p)
names(h)[15] <- "fragvoltotc"
names(h)[16] <- "fragvoltot"

naReplace <- function(h){
  l <- list()
  for(i in seq(names(h))){
    if(class(h[,i])=="character") {l[[i]] <- replace(h[,i], is.na(h[,i]), "NA")} else(l[[i]] <-  h[,i])
  }
  l <- data.frame(l, stringsAsFactors=FALSE)
  names(l) <- names(h)
  return(l)
}

h <- naReplace(h)

# Function
conditional.l.rv.h.summary <- function(x) {
  variable <- unique(x$variable)
  v <- na.omit(x$value) # extract column, from long-formatted input data
  precision <- if(variable == 'phfield') 1 else 0
  ci <- quantile(v, na.rm=TRUE, probs=p) 
  d <- data.frame(min=ci[1], low=ci[2], rv=ci[3], high=ci[4], max=ci[5], stringsAsFactors=FALSE) # combine into DF
  d$range <- with(d, paste("(", paste(round(c(min, low, rv, high, max), precision), collapse=', '), ")", sep="")) # add 'range' column for pretty-printing
  return(d[6])
}
```
## Plot of all pedons in selected set
Pedons that do not have their Std_Latitude and Std_Longitude columns populated in the NASIS Site table are currently not ploted on the map.

```{r plot of pedons and locations, echo=FALSE}
test <- s[complete.cases(c(s$x)),]
test <- s[complete.cases(c(s$y)),]
coordinates(test) <- ~x+y
plot(test, pch=16)
map("county", add=T)
map("state",lwd=2,add=T)
map.axes()
plot(f, label="pedon_id")
```


# Summary of Sites
```{r format site data, echo=FALSE}
# Site information
subset(s, select=c("pedon_id", "taxonname", "tax_subgroup", "part_size_class", "pedon_type", "describer"))

# Summarize surface rock fragments
s.rf <- subset(s, select=c("surface_fgravel", "surface_gravel", "surface_cobbles", "surface_stones", "surface_boulders", "surface_flagstones", "surface_channers"))
s.rf <- replace(s.rf, is.na(s.rf), 0)
s.rf$total_srf <- s.rf$surface_gravel+s.rf$surface_cobbles+s.rf$surface_stones+s.rf$surface_boulders+s.rf$surface_flagstones+s.rf$surface_channers
s.rf$surface_gravel <- s.rf$surface_gravel-s.rf$surface_fgravel
s.m <- melt(s.rf, measure.vars=c("total_srf", "surface_fgravel", "surface_gravel", "surface_cobbles", "surface_stones", "surface_boulders", "surface_channers", "surface_flagstones"))
s.summary <- ddply(s.m, .(variable), .fun=conditional.l.rv.h.summary)
```


## Box plots of surface rock fragments
Graphical five number summary (outliers, 5th, 25th, median, 75th, 95th, outliers)

```{r bwplot for rf, echo=FALSE}
bwplot(variable~value, data=s.m, scales=list(x="free"), xlab="percent")          
```


## Range in characteristics of surface rock fragments
Five number summary (min, 25th, median, 75th, max)(percentiles)

```{r, echo=FALSE, results='asis'}
print(xtable(cast(s.summary, ~variable, value='range')), type="html")
```

# Summary of Pedons
```{r, echo=FALSE, results='asis'}
# Summarize numeric variables by generic horizon
h$gravel <- h$gravel-h$fine_gravel

hp.m <- melt(h, id.vars="genhz", measure.vars=c('clay', 'sand', 'fine_gravel', 'gravel', 'cobbles', 'stones', 'fragvoltot'))
hp.cs <- ddply(hp.m, .(variable, genhz), .fun=conditional.l.rv.h.summary)

hc.m <- melt(h, id.vars="genhz", measure.vars=c('phfield', 'd_value', 'd_chroma', 'm_value', 'm_chroma'))
hc.cs <- ddply(hc.m, .(variable, genhz), .fun=conditional.l.rv.h.summary)

thk <- ddply(h, .(peiid, genhz), summarize, thickness=sum(hzdepb-hzdept))
thk.m <- melt(thk, id.vars="genhz", measure.vars="thickness")
thk.cs <- ddply(thk.m, .(variable, genhz), .fun=conditional.l.rv.h.summary)

hp.c <- cast(hp.cs, genhz ~ variable, value='range')
hc.c <- cast(hc.cs, genhz ~ variable, value='range')
```


## Horizon designations by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$hzname)), digits=0), type="html", width=10) 
```


## Range in characteristics for generic horizons 
Five number summary (min, 25th, median, 75th, max)(percentiles)

```{r, echo=FALSE, results='asis'}
print(xtable(hp.c), type="html")
print(xtable(hc.c), type="html")
```

## Box plots of numeric variables by generic horizon
Graphical five number summary (outliers, 5th, 25th, median, 75th, 95th, outliers)

```{r, echo=FALSE}
bwplot(factor(genhz, levels=levels(h$genhz)[length(levels(h$genhz)):1]) ~ value|variable, data=rbind(hp.m, hc.m, thk.m), scales=list(x="free"))          
```

## Texture by generic horizon
Contigency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$texture_class)), digits=0), type="html")
```

## Tex Mod & Class by generic horizon
Contigency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$texture)), digits=0), type="html")
```

## Dry hue by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$d_hue)), digits=0), type="html")
```

## Moist hue by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$m_hue)), digits=0), type="html")
```

## Effervescence by generic horizon
Contingency table (counts) 

```{r, echo=FALSE, results='asis'}
print(xtable(addmargins(table(h$genhz, h$effervescence)), digits=0), type="html")
```


## Depths and thickness of generic horizons
Five number summary (min, 25th, median, 75th, max)(percentiles)

```{r, echo=FALSE, results='asis'}
genhz.thk <- ddply(h, .(phiid, genhz), summarize, thickness=sum(hzdepb-hzdept))
genhz.m <- melt(h, id.vars="genhz", measure.vars=c('hzdept', 'hzdepb'))
genhz.thk.m <- melt(genhz.thk, id.vars="genhz", measure.vars='thickness')
genhz.m <- rbind(genhz.m, genhz.thk.m)
d.cs <- ddply(genhz.m, .(variable, genhz), .fun=conditional.l.rv.h.summary)
print(xtable(cast(d.cs, genhz ~ variable, value='range'), digits=0), type="html")
```


## Boxplot of generic horizon thicknesses
Graphical five number summary (outliers, 5th, 25th, median, 75th, 95th, outliers)(percentiles)

```{r, echo=FALSE}
bwplot(factor(genhz, levels=levels(h$genhz)[length(levels(h$genhz)):1]) ~ value|variable, data=genhz.m, scales=list(x="free"), xlab="cm", horizontal=T)          
```

## Depths and thickness of diagnostic horizons
Five number summary (min, 25th, median, 75th, max)(percentiles)

```{r, echo=FALSE, results='asis'}
diag.thk <- ddply(f@diagnostic, .(peiid, diag_kind), summarize, thickness=sum(featdepb-featdept))
diag.m <- melt(f@diagnostic, id.vars="diag_kind", measure.vars=c('featdept', 'featdepb'))
diag.thk.m <- melt(diag.thk, id.vars="diag_kind", measure.vars='thickness')
diag.m <- rbind(diag.m, diag.thk.m)
d.cs <- ddply(diag.m, .(variable, diag_kind), .fun=conditional.l.rv.h.summary)
print(xtable(cast(d.cs, diag_kind ~ variable, value='range'), digits=0), type="html")
```

## Boxplot of diagnostic horizon thicknesses
Graphical five number summary (outliers, 5th, 25th, median, 75th, 95th, outliers)(percentiles)

```{r, echo=FALSE}
bwplot(diag_kind ~ value|variable, data=diag.m, scales=list(x="free"), xlab="cm")          
```

