---
title: "Stats for SS Part 1 - Final Project"
author: "Dave White"
date: "February 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Introduction

During FY15 the Las Cruces SSO was contacted by New Mexico State University and NM Department of Transportation, who were interested in dust mitigation for a large playa along interstate 10 near the AZ - NM border. The project area falls within the Hidalgo County Soil Survey (NM023). A majority of the playa was mapped as miscellaneous area "playa" components and  component, Hondale soils (Fine, mixed, superactive, thermic Typic Natrargids). During field reconnaissance in FY14 & 16, the survey crew determined that there was a new component that was previously un-described. This component classifies as a Fine-loamy, mixed, superactive, thermic Sodic Haplocalcids, and there are currently no soil OSD's that fit this classification within MLRA 41. It was determined during a field review that we needed to establish a new soil OSD to capture this component. 

As the majority of my previous work in R focused in spatial data analysis and modeling, I decided that I would use this opportunity to work with the NASIS tabular data in R. This report contains my process of exploratory data analysis, to review the pedon data being used to create the Highlonesome soil OSD and it's relevant range in characteristics. I have also taken this opportunity to use R Markdown to knit an HTML document as my final product. 

*note this document currently has the following options set: knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE). I did this to clean up the output of the final document. I have also commented out (#) specific lines where I was examining the structure of an object.

## Review of pedon data for the creation of Highlonesome soil OSD

Using the fetchNASIS function of the soilDB package to bring in pedon data from NASIS selected set. My selected set contains pedons collected during FY15 and FY16 for the Animas Valley Playa update soil survey project.
```{r fetchNASIS}

library(soilDB)

# get pedons from NASIS
p <- fetchNASIS() # no warning errors when loading pedons

length(p) 
# There are 42 pedons in the selected set

# reviewing the SPC object p
class(p)
#str(p)

# get list of site ids for reference puropsesx
s.id <- p@site$site_id

#print(s.id)
```

## Pedon Locations
Creating a map of the pedon locations
```{r Maps}
library(sp)
library(maps)

# subsetting WGS84 decimal degree corrdinates from p SPC
p.locations <- site(p)[, c('site_id', 'x_std', 'y_std')]

# initialize coordinates in an SPDF
coordinates(p.locations) <- ~ x_std + y_std

# define coordinate system
proj4string(p.locations) <- '+proj=longlat +datum=WGS84'

# set plot margins
par(mar=c(1,1,1,1))

# plot county boundaries for all of NM
map('county', 'new mexico')

# add pedon data locations
points(p.locations, cex = 0.5, pch = 3, col = 'red')

# all of the pedons fall within Hidalgo County, need to zoom in to map

# get extent from p.locations

p.locations@bbox # used this as a rough quide for extent of map

# set up custom xy coordinates for new map extent
xmin <- p.locations@bbox[1,1]-.04
xmax <- p.locations@bbox[1,2]*-1 +.1
ymin <- p.locations@bbox[2,1]-.1
ymax <- p.locations@bbox[2,2]-.3

xlim <- c(xmin, xmax)
ylim <- c(ymin, ymax)

# create new map
map('county', 'New Mexico', xlim=xlim, ylim=ylim)

#add pedon data locations, not the symbol styling
points(p.locations, cex = 1, pch = 3, col = 'red')
box()

# write points to shape file for future use
library(rgdal)

writeOGR(p.locations, dsn = "C:/workspace1", layer = "highlonesome.shp", driver = "ESRI Shapefile", overwrite_layer = TRUE)
```


## Adding Pedon Horizon Lab Results Data
Our office collects pH and EC in the field, using hand held digital meters. The current data structure of NASIS only allows us to populate EC in the pedon horizon lab results table. I realized that the SPC does not contain the pedon horizon lab results. The following was my work around. I have since realized that there is another way to do this utilizing the fetchNASISLabData() function or if using fetchNASIS(lab = TRUE).


```{r adddata}
# We are interested in seeing the range of EC values for the surface horizon, since we populate these values in the pedon horizon lab results, I will need to bring in the data as a csv and join to the SPC

library(readr)
# bringing in csv file containg ec data for every horizon
ec <- read_csv("C:/workspace1/ec.csv")

# inspecting ec object
head(ec)
#str(ec)
#head(p)

# changing column names
names(ec) <- c('peiid', 'phiid', 'name', 'top', 'bottom', 'ec')

# convert to data frame
ec.df <- as.data.frame(ec)

class(ec.df)
head(ec)

# set up new data frame ith only phiid and ec
ec.df <- as.data.frame(ec[, c(2,6)])

# extract horizon data from p
h <- horizons(p)

#str(h)

# add ec data to h 
h1 <- merge(h, ec.df, by = 'phiid')

#head(h1)

# create new SPC p1 from p and replace horizons with h1
p1 <- p
horizons(p1) <- h1
#str(p1)


```


# Review of pedons
Highlonesome is a Fine-loamy, mixed, superactive, thermic, Sodic Haplocalcids soil formed in lacustrine and alluvial deposits of playa lakes found typically around the margins in playa steps.


## Review of Landform data
```{r geomorphreview}

library(Hmisc)# required for plot

# create lf object 
lf <- sort(table(p1$landform_string), decreasing = TRUE)

# plot of landforms
dotchart2(lf, col = 'black', xlim = c(0, max(lf)), cex.labels = 0.75)

# note 3 pedons should be reviewed one has fan piedmont, one has fan skirt, and one is swale, review in ArcGIS and fix in NASIS or remove from selected profiles
head(lf)
#print(lf)
lf.rm <- lf[c(7:9)]
names(lf.rm)


# create an index with landforms to review
lf.idx <- which(p1$landform_string == 'swale' | p1$landform_string == 'fan piedmont' | p1$landform_string == 'fan skirt & playa step')

lf1 <- p1[lf.idx, ]

# site ids to review landforms on
print(lf1$site_id)

#> print(lf1$site_id)
#[1] "2015NM023001" "2015NM023147" "2016NM023017"
```


## Review of Textures
Due to the nature of these soils, rock fragments may be present but at less than 15 percent.
```{r texturereview}
# looking into soil textures
sort(table(p$texture), decreasing=TRUE)

# there are seversal gravelly and very gravelly textured pedons in NASIS, I want to remove them because Highlonesome sould have < 15% gravels

# inspecting data slots of SPC object p
#p$texture
#p$texture_class
#p$fragvoltot

# create an index for gravelly textures
# idx <- which(p$fragvoltot > 15)
#p2 <- p[idx, ]
#p2
# this method was not working for me used subsetProfiles below

p2 <- subsetProfiles(p, h = 'fragvoltot > 15')

table(p2$texture)
#table(p2$fragvoltot)


# plot profiles that have rock fragments
par(mar=c(0,0,2,1))
plot(p2, label='site_id')

# get site_id's for pedons with any horizon that has >% RF
print(p2$site_id)

#> print(p2$site_id)
#[1] "2015NM023074" "2015NM023147" "2015NM023034"
# I changed these to Highlonesome family in NASIS

# I want to remove the profiles with > 15% RF from any further analysis
p1@site$pedon_id

# creating new spc objects without the pedons above
p3 <- p1[2:28]
p4 <- p1[30:36]
p5 <- p1[38:42]

length(p1) 
# original number of pedons

# combining subset spcs into new spc called p1
p1 <- rbind(p3, p4, p5)

length(p1)
# new number of pedons


# creating soil profile plots to look graphically look at the clay contents

par(mar=c(0,0,3,0)) # tighter figure margins
plot(p1[1:19, ], label = 'site_id', color = 'clay')

# note 2016NM023006 dosn't appear to be fine-loamy, review in NASIS
# this pedon is coarse-loamy remove from spc
p1@site$pedon_id

# removing specific record
p1 <- p1[-10, ]

# verifying pedon id's
p1@site$pedon_id


par(mar=c(0,0,3,0)) # tighter figure margins
plot(p1[20:38, ], label = 'site_id', color = 'clay')
```


# Range in Characteristics

## Site Data
```{r shdata}
library(plyr)
library(soilReports)
library(reshape2)
library(knitr)

# get horizon and site level data from the SPC
h <- horizons(p1)
s <- site(p1)
```


## Elevation, Slope, Aspect
```{r elev}
# creating a vector of variables to use below
vars <- c('elev_field', 'slope_field')

# extracting elevation and slope from the site data
elev.slp <- subset(s, select = vars)

# not necessary, but renaiming variables
colnames(elev.slp) <- c("Elevation", "Slope")

# convert data from wide format to long for the plyr package
elev.slp <- melt(elev.slp)

# create a table with the ric
elev.slp.t <- ddply(elev.slp, .(variable), summarize, range = prettySummary(value))

# aspect - have to treat data differently since it is circular data
# creating a vector of which variable to select
vars <- c("aspect_field")

# extracting aspect data from site data
aspect <- subset(s, select = vars)

# renaming column
colnames(aspect) <- c("Aspect")

# converting data from wide format to long
aspect <- melt(aspect)

library(circular)# necessary for circular data
# extracting aspect values and replacing with circular data
aspect$value <- circular(aspect$value, template="geographic", units="degrees", modulo="2pi")

# create table with summarized ric
aspect.t <- ddply(aspect, .(variable), summarize, range = prettySummary(value))

# combined Elev, Slp, and Aspect using the rbind below
kable(rbind(elev.slp.t, aspect.t), caption = "Elevation, Slope, Aspect (min, 25th, median, 75th, max)(n)", align = 'c')
```
 

## Slope Shape
```{r slopeshape}
# inspecting slope shape
s$shapedown
s$shapeacross

# creating a vector of variables
vars <- c('shapedown', 'shapeacross')

# subsetting site data selecting only variables
ss <- subset(s, select = vars)

# inspecting slope shape data table
#ss.t <- table(ss)

# dropping un-used data levels
ss.drop <- droplevels(ss)

# creating data table with slope shape
ss.t2 <- xtabs(~ shapedown + shapeacross, data = ss.drop, drop.unused.levels = TRUE)

# inspecting frequency table
ftable(ss.t2)

kable(ss.t2, caption = "Slope Shape (counts)", align = 'c')
```


## Landform and Parent material
```{r landparent}
# review of gomorphic positions
library(Hmisc)# required for plot

# create lf object 
lf <- sort(table(p1$landform_string), decreasing = TRUE)

# plot of landforms
dotchart2(lf, col = 'black', xlim = c(0, max(lf)), cex.labels = 0.75)

## parent material vs landform

# select variables
vars <-  c('pmkind', 'landform_string')

# subset data
pm.l <- subset(s, select = vars)

# drop levels
pm.l <- droplevels(pm.l)

# create table
pm.l.t <- xtabs(~ pmkind + landform_string, data = pm.l)

# inspecting frequency table
ftable(pm.l.t)

kable(pm.l.t, caption = "Parent Material vs. Landform (counts)", align = 'c')
```


## Drainage
```{r Drainage}
# select variables
vars <- c('drainagecl')

# subset data
drain <- subset(s, select = vars)

# rename cols
colnames(drain) <- c('Drainage Class')

# drop levels
#drain <- droplevels(drain)

# create table
drain.t <- table(drain)

drain.t # inspect table

kable(drain.t, caption = "Drainage Class", align = 'c')
```


## Earth Cover Kind
earth cover kind 1 is on the x-axis and earth cover kind 2 is on the y-axis
```{r coverkind}
# creating a vector of variables
vars <- c('earthcovkind1', 'earthcovkind2')

# subsetting data
eck <- subset(s, select = vars)

#eck #inspecting data element

eck <- droplevels(eck) #removing other levels

# create table
eck.t <- xtabs(~ earthcovkind1 + earthcovkind2, data = eck)

#eck.t #inspecting table

kable(eck.t, caption = "Earth Cover Kind - earth cover kind 1 is on the x-axis and earth cover kind 2 is on the y-axis", align = 'c')
```


## Taxonomy
```{r taxonomy}
# create a vector of variables
vars <- c('pedon_id', 'taxonname', 'taxpartsize', 'taxsubgrp')

# selecting the subset of data
tax <- subset(s, select = vars)

# changing the col names
colnames(tax) <- c('Pedon ID', 'Taxon Name', 'Particle Size Class', 'Subgroup')

kable(tax, caption = 'Review of Pedon Taxonomy', align = 'c')
```


## Diagnostic features
```{r diagfeat}
# extract diagnostic data from spc
d <- p1@diagnostic

#d #inspecting data

d <- d[, 2:4] # removing peiid col

#d #inspecting data

# using the transform function to add thickness
d <- transform(d, 
               thickness = featdepb - featdept,
               featkind = as.character(featkind) 
)

#d #inspecting data

# changing from wide to long format
d <- melt(d, id.vars = 'featkind', measure.vars = c('featdept', 'featdepb', 'thickness'))

#d #inspecting data

# create a table with the ric
d.t <- ddply(d, .(variable, featkind), summarize, range = prettySummary(value))

# dcast used to aggregate data back to feature kind on x axis and depths and thick on y
d.t <- dcast(d.t, featkind ~ variable, value.var = 'range')

kable(d.t, caption = "Diagnostic Features", align = 'c')
```


## Surface Rock Fragments
```{r surfrockfrag}
# a direct copy from the Roecker reports

# grabbing the data from the site object
srf <- s[grepl("surface_", names(s))]

# changing the names
names(srf) <- gsub("surface_", "", names(srf))

# adding total surface fragments
srf <- within(srf, {
  total_srf = gravel + cobbles + stones + boulders + flagstones + channers
  gravel = gravel - fgravel
})

# creating a list of variables to select from
vars <- c("total_srf", "fgravel", "gravel", "cobbles", "stones", "boulders", "channers", "flagstones")

# transforming data from wide to long format
srf.lo <- melt(srf, measure.vars = vars)

# running statistical analysis
srf.5n <- ddply(srf.lo, .(variable), summarize,
                range = prettySummary(value)
)

kable(srf.5n, caption =  "Surface rock fragments (min, 25th, median, 75th, max)(n)", align = 'c')
```


## Generalized Horizon Designations
```{r genhz}

# viewing horizon designations
sort(table(h$hzname), decreasing = TRUE)

# generalized horizon labels
n <- c("An",
       "Bkn1",
       "Bkn2",
       "Cn")
# REGEX rules
r <- c("A",
       "B",
       "2|3|4|kk|^Bknz$",
       "C")

# Compute genhz labels and add to dataset
h$genhz <- generalize.hz(h$hzname, n, r)

p1$genhz <- generalize.hz(p1$hzname, n, r)
```


## Comparison of genhz and hzname pattern matching
```{r genhzpattern}
# select horizon names and gen hz lables
hz_t <- addmargins(table(h$genhz, h$hzname))

# create an index to limit the table size
idx <- pIndex(hz_t, 10)

# plot 10 horizon designations per row
for (i in unique(idx)){
  print(kable(hz_t[, c(idx == i)], align = 'c', digits = 0, caption = "Horizon designations vs generic horizon designations (counts)"))
}
```


## Profile Plots
```{r profileplots}
library(RColorBrewer)

cols <- brewer.pal(n = length(levels(p1$genhz)), name = "Set1") 
hz.names <- levels(p1$genhz)

# assign a color to each generalized horizon label
p1$genhz.soil_color <- cols[match(p1$genhz, hz.names)]

# create an index to limit the number of plots to print
idx <- pIndex(p1, 10)

# plot 10 profiles at a time
for (i in unique(idx)){
  plot(p1[idx == i], name = 'hzname', color = 'genhz.soil_color', label = 'pedon_id')
  title("Soil profile plots")
  legend('bottomleft', legend = hz.names, pt.bg = cols, pch = 22, horiz = TRUE, pt.cex = 2, text.width = 1)
}
```


## General Horizon Depths
```{r genhzdepths}

# basically a direct copy from Roecker reports

# transforming data from wide to long format
genhz.lo <- melt(p1@horizons, id.vars="genhz", measure.vars = c('hzdept', 'hzdepb'))

# creating thickness data from horizon depths
genhz.thk <- ddply(p1@horizons, .(phiid, genhz), summarize, thickness=sum(hzdepb-hzdept))

# transforming data wide to long
genhz.lo2 <- melt(genhz.thk, id.vars = "genhz", measure.vars = 'thickness')

# combining hzn depths and thickness 
genhz.lo <- rbind(genhz.lo, genhz.lo2)

# creating summary statistics
genhz.5n <- ddply(genhz.lo, .(variable, genhz), summarize,
                  range = prettySummary(value)
)

# dcast is changing the structure so that genhz is on the x acis and the variable is on the y
kable(dcast(genhz.5n, genhz ~ variable, value.var = 'range'), align = 'c', caption = "Depths and thickness of generic horizons (min, 25th, median, 75th, max)(n)")
```


## Texture and Rock Frags
```{r texture}
# creating a vector of variable names
vars <- c('clay', 'silt', 'sand', 'total_frags_pct')

# getting data from variables and transforming it from wide to long format
h.tex <- melt(p1@horizons, id.vars = "genhz", measure.vars = vars)

# creating summary statistics
h.tex <- ddply(h.tex, .(variable, genhz), summarize, range = prettySummary(value))

# aggregating data by genetic horizon
h.tex <- dcast(h.tex, genhz ~ variable, value.var = 'range')

kable(h.tex, align = 'c', caption = "Particle size seperates by genetic horizon (min, 25th, median, 75th, max)(n)" )
```


## Texture Class
```{r texclass}
# pulling texture classes as its own table
h.texcl <- h$texcl

# inspecting the frequency table
ftable(h.texcl)

# generating count table of texture classes by genetic horizon
hz_tex <- addmargins(xtabs(~ genhz + texture, data = p1@horizons))

# inspecting texture table
hz_tex

# rearranging cols
kable(hz_tex[, c(4,5,10,9,3,6,2,8,7,1)], caption = "Texture Class by genetic horizon (counts)", align = 'c')



# create soil texture tiangle plots
library(soiltexture)


texture.df <- cbind.data.frame(p1@horizons$genhz, p1@horizons$sand, p1@horizons$silt, p1@horizons$clay)
colnames(texture.df) <- c("genhz", "SAND", "SILT", "CLAY")

# subset by gen hz
h.text.a <- subset(texture.df, texture.df$genhz == 'An')
h.text.b1 <- subset(texture.df, texture.df$genhz == 'Bkn1')
h.text.b2 <- subset(texture.df, texture.df$genhz == 'Bkn2')
h.text.c <- subset(texture.df, texture.df$genhz == 'Cn')


#texture plots
TT.plot( class.sys = "USDA-NCSS.TT",
         tri.data = h.text.a, 
         main = "Soil Texture for An horizons", 
         col = "red", 
         cex = .5, 
         grid.show = FALSE,
         cex.axis = .5,
         cex.lab = .7)

# kept getting an error in teh plot because the sums on some of the horizons were not totalling up to 100
# reviewed nasis data and made changes
#h.tex.b1$sum = h.tex.b1$SAND + h.tex.b1$SILT + h.tex.b1$CLAY

# Texture plot
TT.plot( class.sys = "USDA-NCSS.TT",
         tri.data = h.text.b1, 
         main = "Soil Texture for Bkn1 horizons", 
         col = "red", 
         cex = .5, 
         grid.show = FALSE,
         cex.axis = .5,
         cex.lab = .7)



#texture plot
TT.plot( class.sys = "USDA-NCSS.TT",
         tri.data = h.text.b2, 
         main = "Soil Texture for Bkn2 horizons", 
         col = "red", 
         cex = .5, 
         grid.show = FALSE,
         cex.axis = .5,
         cex.lab = .7)


#texture plot
TT.plot( class.sys = "USDA-NCSS.TT",
         tri.data = h.text.c, 
         main = "Soil Texture for Cn horizons", 
         col = "red", 
         cex = .5, 
         grid.show = FALSE,
         cex.axis = .5,
         cex.lab = .7)



```


## Chemistry
```{r chemistry}
# vector containing which variables are being selected
vars <- c('phfield', 'ec')

# data transformation from wide to long format
h.chem.1 <- melt(p1@horizons, id.vars = "genhz", measure.vars = vars)

# summarize data values by genetic horizon
h.chem <- ddply(h.chem.1, .(variable, genhz), summarize, range=prettySummary(value))

# aggregate data back by genetic horizon
h.chem <- dcast(h.chem, genhz ~ variable, value.var = 'range')

kable(h.chem, align = 'c', caption = "Chemical properties by genetic horizon (min, 25th, median, 75th, max)(n)")

library(lattice)
library(latticeExtra)

# checking out the distribution of ph and ec data by genetic horizon
bwplot(genhz ~ value | variable, data = h.chem.1,
       main = "Box Plots of pH and EC by genetic horizon",
       scales=list(x="free"), axis = axis.grid,
       as.table = TRUE, layout = c(2,1)
       )
```


## Effervescence
```{r effervescence}
# changing name 
p1@horizons$effervescence[p1@horizons$effclass == "very slight"] <- "very.slight"

# storing levels as vector
eff <- c(violent = "violent", strong = "strong", slight = "slight", very.slight = "very slight", none = "none", missing = "missing")

# applying lables as a factor
p1@horizons$effervescence <- eff[p1@horizons$effervescence]
effclass <- factor(p1@horizons$effervescence, levels = eff)

# create table of effervescence class by genetic horizon
effclass.t <- xtabs(~ genhz + effclass, data = p1@horizons, drop.unused.levels = FALSE)

# viewing table
effclass.t

# creating table with ordered cols
kable(addmargins(effclass.t[, c(5,1,2,3,4)]), caption = "Effervescence by genetic horizon counts)", align = 'c')
```


## Salinity
```{r Salinity}
# assigning salinity class based on ec data directly to new saltclass slot

p1@horizons$saltclass <- ifelse(
  p1@horizons$ec < 2, "nonsaline",
  ifelse(
      p1@horizons$ec < 4, "very.slightly",
      ifelse(
          p1@horizons$ec <8, "slightly",
          ifelse(
              p1@horizons$ec < 16, "moderately",
              ifelse(
                  p1@horizons$ec >= 16, "strongly", 
                  "na"
                )
            )
        )
    )
)


# vector containing salt classes
salt <- c(nonsaline = "nonsaline", very.slightly = "very slightly", slightly = "slightly", moderately = "moderately", strongly = "strongly", na = "na")

# applying names to salt class 
p1@horizons$saltclass <- salt[p1@horizons$saltclass]
saltclass <- factor(p1@horizons$saltclass, levels =salt)

# create salt class table gy genetic horizon
saltclass.t <- xtabs(~ genhz + saltclass, data = p1@horizons, drop.unused.levels = FALSE)

# reorder table columns
saltclass.t[ , c(2,5,3,1,4)]

# final table with reorderd cols 
kable(addmargins(saltclass.t[, c(2,5,3,1,4)]), caption = "Salinity Class by genetic horizon (counts)", align = 'c')
```


## Alkalinity (pH Class)
```{r alkalinity}
# converting ph values to class

p1@horizons$phclass <- ifelse(
    p1@horizons$phfield < 6.0, "moderately.acid",
    ifelse(
      p1@horizons$phfield < 6.5, "slightly.acid",
      ifelse(
        p1@horizons$phfield < 7.3, "neutral",
        ifelse(
          p1@horizons$phfield < 7.9, "slightly.alkaline",
          ifelse(
            p1@horizons$phfield < 8.4, "modeately.alkaline",
            ifelse(
              p1@horizons$phfield <= 9, "strongly.alkaline",
              ifelse(
                p1@horizons$phfield > 9, "very.strongly.alkaline",
                "na"
              )
            )
          )
        )
      )
    )
)



# creating a vector of levels
ph <- c(moderately.acid = "moderately acid", slightly.acid = "slightly acid", neutral = "neutral", slightly.alkaline = "slightly alkaline", moderately.alkaline = "moderately alkaline", strongly.alkaline = "strongly alkaline", very.strongly.alkaline = "very strongly alkaline", na = "none")

# applying names
p1@horizons$phclass <- ph[p1@horizons$phclass]
phclass <- factor(p1@horizons$phclass, levels = ph)


kable(addmargins(xtabs(~ genhz + phclass, data = p1@horizons, drop.unused.levels = TRUE)), digits = 0, caption = "pH Class by generic horizon", align = 'c')

```


## Color
```{r color}

# Hue
kable(addmargins(xtabs(~ p1@horizons$genhz + p1@horizons$d_hue, data = p1@horizons, drop.unused.levels = TRUE)), digits = 0, caption = "Dry hue by generic horizon (counts)")

kable(addmargins(xtabs(~ genhz + m_hue, data = p1@horizons, drop.unused.levels = TRUE)), digits = 0, caption = "Moist hue by generic horizon (counts)")

# Value and Chroma
vars <- c('d_value', 'd_chroma', 'm_value', 'm_chroma')

h.col <- melt(p1@horizons, id.vars = "genhz", measure.vars = vars)
h.col <- ddply(h.col, .(variable, genhz), summarize, range=prettySummary(value))
h.col <- dcast(h.col, genhz ~ variable, value.var = 'range')

kable(h.col, align = 'c', caption = "Color(Value and Chroma) by genetic horizon (min, 25th, median, 75th, max)(n)" )

```
